\begin{abstract}
Artificial intelligence is increasingly used to construct predictive and
simulation-based models of complex scientific systems. However, model
evaluation is commonly reduced to predictive accuracy, even when the intended
claim concerns reproducibility, structural fidelity, or mechanism. This paper
introduces MouseBrainBench, an engineering framework for verification and
validation of partial digital models of the mouse brain. Its contribution to
artificial intelligence is a claim-aware benchmark architecture that represents
prediction, reproducibility, topology specificity, directed identifiability,
structure--function agreement, and computational cost as non-interchangeable
evidence blocks. A Mechanistic Identifiability Score provides descriptive
diagnostics, while a conjunctive decision rule prevents high performance in one
block from compensating for missing evidence in another.

The framework is evaluated with controlled simulations and three families of
public data. Synthetic systems confirm that the decision rule rejects common
drive, topology without temporal direction, and direction without
topology-specific prediction. Allen Visual Behavior Neuropixels provides a real
negative control: median cross-mouse and split-half correlations reach 0.891 and
0.990, but topology-specific and directed-identifiability criteria fail. Static
and Dynamic Sensorium demonstrate predictive gains without corresponding
mechanistic evidence. MICRONS provides an external structure--function
validation case. Across a discovery cohort and two non-overlapping hold-outs
containing 2991 co-registered units and 5943 connected directed pairs,
connected pairs exhibit greater readout-location similarity than distance- and
degree-matched controls. Distance-matched effects are 0.0215, 0.0210, and
0.0179, with positive unit-cluster bootstrap intervals in all cohorts.

MouseBrainBench does not claim a complete or causal digital brain. It provides
a reproducible engineering layer for deciding which claims about AI-based
scientific models are supported, partially supported, or blocked by the
available evidence.
\end{abstract}
