\section{Conclusions}
\label{sec:conclusions}

This paper presented MouseBrainBench, a claim-aware verification and validation
framework for \acs{ai} models used as partial digital representations of the
mouse brain. The proposal tackles a relevant problem: predictive performance,
reproducibility, structural fidelity, and mechanistic identifiability are often
measured separately but interpreted as if they were interchangeable. To avoid
this situation, MouseBrainBench organises validation through explicit evidence
blocks, matched controls, and non-compensatory claim gates.

The Mechanistic Identifiability Score provides a diagnostic summary of the
available evidence. However, the final claim is not decided by a weighted
average. A claim passes only when all the required evidence blocks pass. The
synthetic experiments confirm this behaviour by rejecting common drive, topology
without direction, and direction without topology-specific prediction, even when
some partial evidence is strong. This fact is the main methodological
contribution of the work, as it turns validation assumptions into executable and
falsifiable contracts.

The real-data experiments produced complementary outcomes. Allen Visual Behavior
Neuropixels showed strong reproducibility, but did not identify the proposed
topology or direction. Static and Dynamic Sensorium provided predictive gains,
but the evidence was not sufficient to support a mechanistic or causal claim.
MICRONS provided the main positive empirical case: across a discovery cohort and two
non-overlapping hold-out windows from the same resource, connected directed
pairs showed greater readout-location similarity than distance- and
degree-matched controls, with positive unit-cluster bootstrap stability intervals.

It can be concluded that MouseBrainBench is useful as a claim-audit framework
for partial mouse-brain digital models. The current evidence supports a bounded
local observational structure--function claim in MICRONS and demonstrates that
several tempting broader claims must remain blocked. In particular, the paper
does not validate a complete mouse-brain digital twin, a causal mechanism, a
behavioural twin, or a state-of-the-art Sensorium predictor.

Future work should apply the framework to external simulators, including
biophysically detailed models, neural-mass models, and emerging digital-brain
proposals. It is also necessary to extend the framework towards intervention
experiments, broader subject-level replication, stronger uncertainty
quantification, larger Monte Carlo operating-characteristic studies, and direct
comparisons with conventional predictive-only evaluation. These extensions should preserve the central principle of the
proposal: every model output must be interpreted only according to the evidence
that has actually passed the declared claim gate.
