\section{Datasets and roles}

The benchmark uses several datasets, but each one has a different role in the
proposal. This decision is fundamental because not all datasets can support the
same type of claim. Some resources are useful to evaluate prediction, others to
evaluate reproducibility, and others to evaluate local structure-function
relationships.

\subsection{Synthetic known-truth cases}

Four synthetic cases are included to validate the Mechanistic Identifiability
Score under controlled conditions. These cases are directed truth, common drive,
structure without temporal direction, and timing without region-specific
topology. They do not provide biological evidence. Their role is to verify that
the score accepts the designed positive case and rejects typical failure modes.

\subsection{Allen Visual Behavior Neuropixels}

Allen Visual Behavior Neuropixels is used as a negative
mechanistic-identifiability control. The analysis includes 20 successful
sessions in the confirmation phase, 20 sessions and 20 unique mice in the
anatomical-connectivity test, and 21 sessions with 19 unique mice in the
directed-identifiability test. This dataset is expected to provide reproducible
neural information, but it should not be sufficient to support a directed
anatomical mechanism under the current evidence gates.

\subsection{Sensorium static}

Sensorium static is used as an auxiliary predictive and reliability case. The
local package includes seven validation mice and five repeated-test mice. Across
the repeated-test mice, median reliability is 0.642 and median best predictive
correlation is 0.341. Therefore, this layer is useful to analyse prediction and
repeat reliability, but it is not used as direct causal evidence.

\subsection{Dynamic Sensorium}

Dynamic Sensorium is used as a temporal predictive stress test. The current
package includes five current public mice and five legacy out-of-distribution
mice. Transparent baselines include mean response, summary features, temporal
filterbank, temporal singular value decomposition, random features, a compact
PyTorch multilayer perceptron, and a bounded official Sensorium model family.
The official bounded baseline is integrated and evaluated, but it is not treated
as a state-of-the-art or leaderboard-equivalent result.

\subsection{MICRONS CAVE}

MICRONS CAVE is the main empirical evidence layer of the manuscript. The Q1
package contains three non-overlapping windows: one discovery cohort and two
offset hold-out cohorts. Together, these cohorts cover 2991 co-registered
units, 6575 loaded synapses, and 5943 connected directed pairs. MICRONS is used
because it allows anatomical connectivity and functional readout properties to
be analysed in the same local cortical resource.
