\section{Experimental methodology}

\subsection{Study design and claim hierarchy}

The study was designed as a methodological validation rather than as a
competition between datasets or model architectures. Each dataset was assigned
a role before analysis according to the observations it contains. Synthetic
systems provide known truth; Allen Visual Behavior Neuropixels provides repeated
population activity without individual synaptic connectivity; Sensorium
provides standardized predictive targets; and MICRONS provides co-registered
local anatomy and functional descriptors. These resources are therefore not
treated as interchangeable samples from one population.

A benchmark instance contains five elements: a bounded claim, observable
quantities, required evidence blocks, decision criteria, and provenance. The
claim is formulated before metric selection. For example, the claim that a
specific directed topology is identifiable requires reproducibility, superiority
over plausible alternative topologies, and a resolvable directional signature.
By contrast, the claim that connectivity is locally associated with a functional
descriptor requires co-registered structure and function, controls for major
anatomical confounds, multiplicity correction, and independent replication. It
does not require temporal direction because it is explicitly observational.

The analysis follows a sequence of increasingly realistic tests. First, the
decision rule is evaluated under known truth. Second, a real negative control
tests whether reproducibility is incorrectly promoted to mechanism. Third,
predictive benchmarks test whether model accuracy is kept separate from
anatomical evidence. Finally, the MICRONS analysis tests whether the framework
can admit a positive but bounded structure--function claim. The publication
freeze is generated only after all component artifacts are available.

\subsection{Mechanistic Identifiability Score}

Let $C$ be a scientific claim and let $B(C)=\{B_1,\ldots,B_m\}$ be its
required evidence blocks. Each block $B_i$ contains criteria
$c_{ij}=(x_{ij},t_{ij},d_{ij})$, where $x_{ij}$ is an observed statistic,
$t_{ij}$ is a fixed threshold, and $d_{ij}$ specifies whether the statistic
must be greater than, greater than or equal to, less than, or less than or equal
to that threshold. Criterion passage is
\[
p_{ij}=\mathbb{I}\left[x_{ij}\ d_{ij}\ t_{ij}\right].
\]
For criteria requiring a larger value, the descriptive normalized score is
$s_{ij}=\min(1,\max(0,x_{ij}/t_{ij}))$. For criteria requiring a smaller
value, the inverse ratio is used. Zero-threshold criteria receive one when they
pass and zero otherwise.

The score of block $i$ is the arithmetic mean
\[
S_i=\frac{1}{n_i}\sum_{j=1}^{n_i}s_{ij},
\]
and the displayed MIS value is $M=m^{-1}\sum_i S_i$. However, a block passes
only when all of its criteria pass, $P_i=\bigwedge_j p_{ij}$, and the
mechanistic decision is
\[
P_{\mathrm{MIS}}=\bigwedge_{i=1}^{m}P_i.
\]
The continuous value $M$ is diagnostic; it does not override the conjunctive
decision. This prevents strong reproducibility from compensating for absent
topology specificity, or strong prediction from compensating for missing
directional evidence.

The canonical implementation contains three blocks. Reproducibility assesses
whether the empirical target is stable across animals, sessions, or independent
partitions. Topology specificity assesses whether the proposed structure
outperforms disconnected, permuted, transposed, or otherwise matched
alternatives. Directed identifiability assesses whether latency, lead--lag
organization, or an equivalent directional signature is resolved reliably. A
dataset-specific audit can add observational structure--function criteria, but
those criteria are not relabelled as causal directionality.

\subsection{Synthetic known-truth systems}

Each synthetic case contains 24 sessions, six regions, and 24 time samples. A
regional template is generated from Gaussian temporal profiles. Directed
latency is introduced by shifting the profile centre by 0.08 time units per
region. Region-specific structure is introduced through systematic changes in
amplitude and width. Session noise has standard deviation 0.08; independent
odd/even perturbations used for split-half evaluation have standard deviation
0.04. Fixed random seeds 11, 23, 31, and 43 identify the four cases.

Reproducibility is measured as the median correlation between individual
sessions and their mean template, median odd--even correlation, and the fraction
of sessions exceeding the 95th percentile of 200 within-session region
permutations. Topology specificity compares the candidate prediction with 100
region-permuted predictions and a disconnected template. Directionality is
measured as the fraction of regional pairs with distinct peak latencies and the
equivalent nonzero lead--lag fraction.

The four systems isolate complementary conditions. The directed-truth case
contains all required information. The common-drive case is reproducible but has
neither region-specific topology nor direction. The topology-without-direction
case contains structured regional amplitudes but simultaneous temporal peaks.
The direction-without-topology case contains staggered peaks while its candidate
prediction is intentionally no more topology-specific than the disconnected
alternative. Before execution, the expected decision was that only the
directed-truth case should pass.

\subsection{Allen Visual Behavior Neuropixels negative control}

Allen analyses use exported population-response artifacts from Visual Behavior
Neuropixels sessions. The reproducibility phase contains 20 successful sessions.
The anatomical test contains 20 sessions from 20 unique mice, while the
directional test contains 21 sessions from 19 mice. Session inclusion, regional
aggregation, time windows, and preprocessing are fixed in the corresponding
configuration and exported with the result artifacts.

The reproducibility block requires median cross-mouse correlation greater than
0.50, median split-half correlation greater than 0.70, and at least half of
sessions above their own 95th-percentile null. The topology block requires an
Allen-informed advantage greater than 0.05 over both the median permutation and
the disconnected model, at least 95\% of permutations outperformed, and a
positive lower bound for the paired bootstrap interval. These thresholds ask
whether anatomical organization contributes information beyond generic regional
dynamics.

The directional block uses six criteria. Median cross-mouse latency Kendall
correlation must exceed 0.30, median split-half latency correlation must exceed
0.50, and at least half the sessions must exceed their own latency null. At least
half the regional pairs must have resolvable latency, at least half the sessions
must exceed the lead--lag null, and the median nonzero lead--lag pair fraction
must be at least 0.25. Failure of this block does not imply absence of biological
directionality; it means that the selected target does not resolve it under the
specified procedure.

\subsection{Static Sensorium evaluation}

Static Sensorium evaluation uses seven validation mice and five mice with
repeated test stimuli. The response of each neuron is compared with transparent
predictors: a stimulus-independent mean response, ridge regression over stimulus
features, and ridge regression over stimulus and behavioural context. A
scrambled-stimulus control preserves response statistics while breaking the
stimulus--response assignment. Pearson correlation is computed on held-out
responses and summarized at the mouse level.

Repeated stimuli provide an empirical reliability estimate. Reliability is not
imputed for tiers without repetitions; those tiers remain informative for
prediction but cannot pass a repeat-reliability gate. The topographic constraint
tests whether similarity in response descriptors varies systematically with
recorded spatial organization. Its Spearman statistic and effect over a matched
null are reported separately from prediction because a topographic association
does not establish synaptic connectivity or causal direction.

The static benchmark is deliberately not scored as a binary model competition.
It asks whether stimulus-aware models beat mean and scrambled controls, whether
the response is repeatable where repetitions exist, and whether partial
topographic evidence is present. Mechanistic passage additionally requires
evidence unavailable in this configuration and is therefore expected to remain
negative.

\subsection{Dynamic Sensorium predictive stress test}

Dynamic Sensorium uses five current public mice and five legacy
out-of-distribution mice. Seven model families are represented in the comparator:
mean response, summary features, temporal filterbank, temporal singular-value
decomposition, random features, a compact PyTorch multilayer perceptron, and a
bounded model instantiated through the official Sensorium loader and model
factory. Predictions are paired by mouse to avoid interpreting between-animal
differences as model effects.

The transparent temporal models provide intermediate complexity between a mean
response and a neural network. The filterbank represents finite temporal
kernels. Temporal singular-value decomposition represents a low-rank temporal
subspace. The random-feature model tests whether nonlinear expansion alone
explains gains, and the multilayer perceptron tests a compact learned nonlinear
mapping. For each mouse, the comparator reports correlation, the winning model,
and paired differences between model families.

The official architecture was trained under a bounded local budget on Apple
Metal Performance Shaders. It verifies that the official software stack,
dataloader, model factory, training, and evaluation can execute in the
MouseBrainBench environment. It does not reproduce the published training
budget or leaderboard protocol and is therefore excluded from state-of-the-art
claims. This distinction is encoded in the audit rather than left as a narrative
qualification.

\subsection{MICRONS data acquisition and cohort construction}

MICRONS data are queried from the public \texttt{minnie65\_public} CAVE stack.
Functional descriptors come from
\texttt{digital\_twin\_properties\_bcm\_coreg\_v4}; structural edges are
queried at materialization version 1507. The acquisition artifact stores the
datastack, table, materialization, query limits, offsets, source paths, and
result counts. CAVE materializations provide a time-locked annotation state
despite continued proofreading \citep{dorkenwald2025cave}. A CAVE token and
acceptance of the service terms are required to reproduce remote queries.

Three non-overlapping ordered windows define discovery and replication. Offset
0 provides 1000 usable units, offset 1000 provides 992, and offset 2000 provides
999. The windows were not pooled before evaluation. Structural queries yielded
2267, 2161, and 2147 synapses, corresponding to 2095, 1926, and 1922 unique
connected directed pairs. All possible directed non-self pairs form the
candidate universe in each cohort.

The primary endpoint is the similarity of the digital-twin readout-location
descriptor for connected pairs. This endpoint was selected after the expanded
aggregate analysis failed its distance-matched gate and was then fixed for the
replication package. The discovery and hold-out artifacts are retained
separately. Additional functional descriptors and anatomical strata are treated
as secondary analyses and are not promoted to the manuscript's primary claim.

\subsection{Matched controls and statistical inference}

For each connected directed pair, three comparator families are constructed.
Random non-connected pairs estimate the unconditioned contrast. Distance-matched
non-connected pairs control for spatial separation because connection
probability and functional organization both vary with distance. Degree-matched
non-connected pairs control for unequal representation of neurons with many
incoming or outgoing edges. The reported effect, $\Delta$, is the connected
pair similarity minus the corresponding control similarity.

Tests are one-sided for the prespecified hypothesis that connected pairs are
more similar than non-connected controls. The Benjamini--Hochberg procedure
controls the false discovery rate over the family of tests within each
stratified analysis. The primary endpoint is considered confirmed only if both
distance- and degree-matched effects are positive after correction in discovery
and in each hold-out cohort.

Pairwise observations are dependent because one neuron contributes to many
pairs. Confidence intervals therefore use a unit-cluster bootstrap rather than
a naive pair bootstrap. Each of 300 samples resamples units with replacement,
reconstructs the eligible pair set, and recomputes matched effects. Seed 71 is
fixed in the reproducibility package. The 2.5th and 97.5th percentiles of usable replicates
form the 95\% interval.

\subsection{Sensitivity, provenance, and publication freeze}

Threshold sensitivity is evaluated without reselecting outcomes. Allen MIS
criteria are multiplied by 0.75, 1.00, and 1.25. Static Sensorium reliability
thresholds of 0.50, 0.60, and 0.70 are crossed with topographic thresholds of
0.05, 0.10, and 0.15. Dynamic Sensorium paired gains are examined at minimum
effects of 0, 0.002, 0.005, and 0.01. These checks determine whether the
qualitative interpretation depends on a single boundary.

The Python package separates loaders, typed schemas, connectivity, dynamics,
simulation, validation, baseline adapters, and visualization. Benchmark commands
write JSON and Markdown summaries containing software version, Git revision,
inputs, sample counts, criteria, and decisions. Automated tests cover criterion
direction, normalization, conjunctive block semantics, known-truth cases,
dataset adapters, and publication gates.

The publication-freeze command reads sealed artifacts and emits two explicit
lists: claims supported by the current evidence and claims that remain blocked.
It does not recalculate primary statistics or permit manual substitution of
headline values. This design keeps manuscript language synchronized with the
reproducible evidence package and makes any future change in claim status
auditable.

\subsection{Ethics and use of public animal data}

This study performs secondary computational analyses of publicly released data
and introduces no new animal procedures. Animal acquisition, experimental
protocols, welfare oversight, and institutional approvals are described in the
original Allen, Sensorium, Dynamic Sensorium, and MICRONS publications cited in
this manuscript. MouseBrainBench operates on released recordings, metadata, and
derived structural resources under their applicable access conditions. No
personally identifiable human data are used.
