\section{Proposed framework}

MouseBrainBench is a claim-aware verification and validation framework for
partial digital models of the mouse brain. It is not a neuronal simulator and
does not replace domain infrastructures such as BMTK, SONATA, The Virtual
Brain, Sensorium, or CAVE. These systems supply models, predictions, or data.
MouseBrainBench supplies the engineering contract that links their outputs to a
bounded scientific claim.

\subsection{Problem formulation}

Let $\mathcal{M}$ denote a computational model, $\mathcal{D}$ an evaluation
dataset, and $C$ a claim made about their relationship. Conventional
benchmarking usually estimates a predictive utility
$U(\mathcal{M},\mathcal{D})$ and ranks models using that scalar. This is
appropriate when $C$ is limited to held-out prediction under the same
evaluation protocol. It is insufficient when $C$ states that $\mathcal{M}$
recovers anatomical organization, directed dynamics, or a biological mechanism.
Several models may be predictively equivalent while encoding incompatible
internal structures.

MouseBrainBench represents a benchmark task as
\[
\mathcal{T} =
\left(C,\mathcal{O},\mathcal{B},\mathcal{N},\mathcal{R},\mathcal{P}\right),
\]
where $\mathcal{O}$ is the set of observable quantities, $\mathcal{B}$ contains
the evidence blocks required by the claim, $\mathcal{N}$ defines null and
alternative models, $\mathcal{R}$ contains decision rules, and $\mathcal{P}$
records provenance. The task is valid only when every quantity needed by
$\mathcal{B}$ is observable in $\mathcal{D}$. Missing evidence is not imputed
from predictive performance.

This formulation makes the claim an input to evaluation rather than a
post-hoc interpretation. A Sensorium predictor can be accepted as a functional
surrogate while being rejected as an anatomical explanation. A local MICRONS
association can pass structure--function controls while remaining insufficient
for a causal or whole-brain claim. The same numerical result can therefore have
different admissible interpretations depending on $C$.

\subsection{Evidence blocks and non-compensatory decision}

An evidence block contains criteria that answer one validation question. The
canonical configuration uses reproducibility, topology specificity, and
directed identifiability. Predictive tasks additionally report held-out
accuracy, out-of-distribution performance, and empirical reliability.
Structure--function tasks introduce matched anatomical controls and independent
replication. Computational cost is recorded as an engineering property but does
not substitute for biological evidence.

For criterion $j$ in block $i$, let $x_{ij}$ be the observed value,
$t_{ij}$ its fixed threshold, and $d_{ij}$ the comparison direction. Criterion
passage is
\[
p_{ij}=\mathbb{I}\!\left[x_{ij}\ d_{ij}\ t_{ij}\right].
\]
For a criterion requiring a value larger than its threshold, the normalized
diagnostic score is
\[
s_{ij}=\min\!\left(1,\max\!\left(0,\frac{x_{ij}}{t_{ij}}\right)\right).
\]
The inverse ratio is used when smaller values provide stronger evidence.
Zero-threshold criteria receive one when they pass and zero otherwise.

The descriptive score of evidence block $i$ is
\[
S_i=\frac{1}{n_i}\sum_{j=1}^{n_i}s_{ij},
\]
whereas block passage is $P_i=\bigwedge_j p_{ij}$. For $m$ blocks, the reported
Mechanistic Identifiability Score (MIS) is
\[
M=\frac{1}{m}\sum_{i=1}^{m}S_i,
\]
and the claim decision is
\[
P_{\mathrm{MIS}}=\bigwedge_{i=1}^{m}P_i.
\]

The separation between $M$ and $P_{\mathrm{MIS}}$ is intentional. $M$ helps
locate weak blocks and compare sensitivity scenarios, but it is not an additive
currency. Reproducibility cannot compensate for a topology that performs like a
permuted graph. Directed timing cannot compensate for failure to distinguish
the proposed regional organization. This non-compensatory behavior differentiates
the framework from a weighted average in which unrelated evidence can cancel
out.

\subsection{Claim classes}

MouseBrainBench distinguishes four claim classes used in this study. A
\emph{predictive claim} states that a model improves held-out neural-response
prediction over specified baselines. A \emph{reproducibility claim} states that
the target or effect is stable across repeated stimuli, sessions, animals, or
partitions. A \emph{structure--function claim} states that a structural relation
is associated with a functional descriptor after declared anatomical controls.
A \emph{mechanistic-identifiability claim} additionally requires that the
proposed topology and direction can be distinguished from plausible
alternatives.

These classes form a partial order rather than a linear quality scale.
Mechanistic identifiability requires reproducible observations, but useful
prediction does not automatically imply structural fidelity. A
structure--function association can be stronger than an uncontrolled
correlation and still weaker than intervention-supported causality. The audit
preserves these intermediate outcomes instead of forcing every benchmark into a
binary useful/useless label.

The framework also defines blocked claim classes. Whole-brain digital-twin
claims require evidence across brain-wide structure, dynamics, calibration, and
perturbation that none of the local resources used here jointly provide.
Behavioural-twin claims require closed-loop or task-level validation.
Consciousness claims have no operational target in the present system. Blocking
these statements is a validation result, not a software limitation disguised as
future functionality.

\subsection{Software architecture}

The implementation follows a layered architecture. The data layer contains
source-specific loaders, typed schemas, cache policies, and provenance records.
The representation layer maps regions, neurons, functional units, directed
edges, stimuli, and responses into stable Python objects. The model layer
contains transparent dynamics and adapters for external predictors. The
validation layer implements reliability, anatomical, functional, perturbation,
cost, and MIS metrics. The orchestration layer executes benchmark plans and
freezes their claim status.

Adapters isolate external ecosystems from the benchmark core. Allen exporters
convert sessions into normalized analysis artifacts. Sensorium adapters expose
static and dynamic predictors through a common response contract. CAVE adapters
query materialized MICRONS tables and preserve datastack and materialization
identifiers. BMTK and TVB adapters are reserved for external simulation outputs;
the framework does not reimplement their numerical engines.

Each benchmark command writes a machine-readable JSON artifact and a
human-readable Markdown summary. Artifacts contain the software version, Git
revision, configuration, source identifiers, cohort size, random seed, metrics,
thresholds, block decisions, and final interpretation. A result can therefore
be traced from a manuscript statement to the exact data and code state that
generated it.

\subsection{Null models and controls}

Null models are part of the task definition. Region permutations test whether a
candidate organization contributes more than label assignment. Disconnected
models test whether interactions add information beyond independent regional
dynamics. Scrambled-stimulus controls break stimulus--response correspondence
while preserving marginal statistics. Distance-matched MICRONS controls account
for spatial embedding, and degree-matched controls account for unequal
representation of highly connected neurons.

Controls are selected according to the confound addressed by the claim. Passing
a weak random null does not imply passage of a stronger matched null. This
principle is visible in the MICRONS experiments: an aggregate effect passes
random and degree controls but fails distance matching, forcing the claim to
remain negative until a stratified endpoint replicates under both controls.

\subsection{Publication freeze and traceability}

The final engineering component is the publication freeze. It reads sealed
benchmark artifacts and produces explicit lists of admitted, partially
supported, and blocked claims. It does not recompute results or permit manual
replacement of headline metrics. Manuscript assertions are generated from this
contract, reducing the risk that exploratory findings are promoted during
writing.

The freeze also reports incomplete baselines and external dependencies. For
example, successful execution of a bounded official Sensorium model is recorded,
but the state-of-the-art claim remains blocked because the published training
budget and leaderboard protocol were not reproduced. This distinction allows
engineering integration to be documented without treating it as empirical
equivalence.

\subsection{Evidential role of each dataset}

No available resource contains every observation required for a complete
digital-twin claim. Synthetic systems test decision logic under known truth.
Allen provides standardized, multi-animal neural activity and a negative
mechanistic control. Static and Dynamic Sensorium provide modern predictive
tasks. MICRONS provides local co-registered structure and function. Their roles
are complementary rather than interchangeable.

Table~\ref{tab:dataset-roles} summarizes this design. A negative MIS result for
Allen or Sensorium does not imply low dataset quality. It states that the
particular observations used here do not identify the stronger mechanism. The
MICRONS result is positive only for a local observational association and does
not upgrade the other datasets or the complete system to a digital twin.

\input{tables/dataset_roles}
