\section{Introduction}
\label{sec:introduction}

Computational representations of biological systems have become a key point in
the interaction between \ac{ai} and neuroscience. In the case of the laboratory
mouse (\textit{Mus musculus}), several public infrastructures provide standardised
neurophysiological recordings, visual-response benchmarks, anatomical atlases,
and electron-microscopy reconstructions. These resources allow researchers to
analyse the same biological system from different perspectives, ranging from
population activity recorded across brain regions to synaptic connectivity in a
restricted cortical volume. Therefore, they also facilitate the development of
models that reproduce selected observations with increasing accuracy.

However, this progress also generates a relevant validation problem. A model
that predicts neural responses is not necessarily a model that identifies the
circuit producing them. A response can be stable across animals while remaining
compatible with several alternative topologies. Similarly, a model can preserve
an anatomical graph without demonstrating that its directed dynamics are
identified by the available observations. Notice that predictive performance,
empirical reliability, anatomical plausibility, and causal explanation answer
different scientific questions. Thus, a single correlation coefficient is not
sufficient to decide whether a model supports a mechanistic claim.

This distinction is especially important when digital-twin terminology is used.
A digital twin normally implies correspondence with a specific physical entity,
repeated calibration, and the ability to reproduce relevant states under
meaningful interventions. Current mouse-brain resources do not jointly provide
complete synaptic connectivity, full-brain dynamics, behaviour, and
interventional validation for an individual animal. \ac{microns} offers
exceptional synaptic and functional detail, but only for a local visual-cortical
volume \citep{bae2025microns}. Allen datasets provide broad and standardised
neural recordings, but not the individual synaptic graph that generated each
response \citep{siegle2021survey}. Consequently,
treating either resource as a complete mouse-brain twin would exceed the
available evidence.

Existing software and benchmark infrastructures already solve relevant parts of
this problem. \ac{bmtk} provides tools for constructing and simulating multiscale
neural networks \citep{dai2020bmtk}. \ac{tvb} supports connectome-based
population dynamics \citep{tvb2013}. Sensorium provides common predictive tasks
for mouse visual cortex \citep{willeke2022sensorium,turishcheva2023dynamic}.
For this reason, the main gap is not the absence of another simulator or another
visual-response leaderboard. The unresolved point is the audit between a metric
and the biological claim attached to it: what evidence is required before a
reproducible prediction can be interpreted as support for a specific mechanism?

In accordance with this objective, this paper presents MouseBrainBench, a
claim-aware validation framework for partial mouse-brain digital models. It is
not proposed as a complete brain simulator and it does not claim to instantiate
a full digital twin. The framework assigns public datasets to explicit
evidential roles, evaluates them using matched null models and hold-out checks,
and records which interpretations are admitted or blocked. Its Mechanistic
Identifiability Score (MIS) is not a conventional weighted ranking. It is a
conjunctive decision procedure over non-interchangeable evidence blocks.

From the perspective of Engineering Applications of Artificial Intelligence, MouseBrainBench is framed as an AI engineering contribution rather than as a new biological simulator. The AI contribution is the executable claim gate: a non-compensatory validation layer that turns qualitative model claims into auditable decisions. The engineering application is the validation of partial neurocomputational digital models before they are used to support scientific or technical interpretation. This formulation also provides a measurable advantage over conventional predictive evaluation, because it identifies cases in which a high predictive or reproducibility score would incorrectly promote an unsupported mechanistic claim.

The proposal is evaluated through positive and negative cases. Four synthetic
systems are used to test whether the MIS recovers known truth and rejects
common-drive or incomplete mechanisms. Allen Visual Behavior Neuropixels is used
to verify that the framework can keep a highly reproducible empirical target
while rejecting unsupported topology and direction claims. Static and Dynamic
Sensorium are used to preserve predictive evidence without promoting it to
causality. Finally, MICRONS is used to test a discovery-fixed local
structure--function endpoint against distance- and degree-matched nulls in a
discovery cohort and two non-overlapping hold-outs. This sequence evaluates the
audit procedure itself, not a collection of unrelated dataset demonstrations.

The rest of the article is organised as follows. Section~\ref{sec:contributions}
details the research contributions of the proposal. Section~\ref{sec:related-work}
reviews the most relevant predictive, simulation, and connectomic approaches.
Section~\ref{sec:framework} describes the MouseBrainBench framework and its
main components. Section~\ref{sec:evaluation} presents the experimental
evaluation and the achieved results. Section~\ref{sec:discussion} discusses the
scope and limitations of the proposal. Finally, Section~\ref{sec:conclusions}
concludes the paper and provides future guidelines.
