\section{Limitations and future guidelines}

The proposal presents several limitations that must be considered before any
submission. First, the MICRONS result is observational. It does not establish
causality, synaptic necessity, or behavioural function. The analysis is also
local to the available cortical volume and cannot be generalized to the whole
mouse brain without additional evidence.

Second, the Allen Visual Behavior Neuropixels analysis is used as a
negative-control layer and currently focuses on a controlled subset rather than
on every available public session. This decision is acceptable for the current
claim because Allen is not the main positive result, but a future version should
expand this layer to increase coverage.

Third, the Sensorium official baseline is a bounded local
training/evaluation artifact. It proves that the official ecosystem can be
executed inside MouseBrainBench, but it is not equivalent to a leaderboard or
state-of-the-art Sensorium result. Therefore, no Sensorium SOTA claim is made.

Finally, MouseBrainBench does not simulate consciousness, behaviour, or a
complete mouse brain. The framework audits partial digital models and the claims
associated with them. Future research will focus on extending the number of
validated datasets, adding perturbation-based evidence when public data allow
it, connecting external simulators through stable adapters, and improving the
formal definition of mechanistic identifiability under different biological
scales.
