\section{Methods}

\subsection{Software implementation}

MouseBrainBench is implemented as a Python package with reproducible benchmark
scripts, configuration files, and tracked result artifacts. The implementation
produces JSON and Markdown summaries for the most relevant analyses. These
artifacts are used to generate a publication freeze that records which claims
are allowed and which claims are blocked.

\subsection{Claim-audit procedure}

The claim-audit procedure receives the output of each benchmark layer and maps
it to a set of possible interpretations. A claim is accepted only when all the
required evidence gates pass. If one relevant gate fails, the claim is blocked
even if other metrics are positive. This behaviour is especially important for
distinguishing predictive models from mechanistic explanations.

\subsection{Mechanistic Identifiability Score}

The Mechanistic Identifiability Score evaluates reproducibility, topology
specificity, directed identifiability, and, when available, structure-function or
perturbation constraints. The score is conservative and conjunctive. Thus, a
model can be accepted as a useful predictor and rejected as a mechanistic model
in the same audit.

\subsection{Allen analysis}

Allen Visual Behavior Neuropixels analyses evaluate whether reproducible neural
signatures align with anatomical and directed-identifiability constraints. The
final decision is negative because topology specificity and directionality fail
despite strong reproducibility. This layer is used to verify that the framework
does not overclaim from reproducible activity alone.

\subsection{Sensorium analysis}

Sensorium static analyses quantify reliability and predictive improvement over
simple controls. Dynamic Sensorium analyses compare mean response, summary
features, temporal filterbank, temporal singular value decomposition, random
features, compact neural-network models, and a bounded official Sensorium
baseline. The official bounded baseline uses the official Sensorium loader and
model factory under a local bounded training budget. Therefore, it is treated as
an integration control rather than a Q1-qualified leaderboard baseline.

\subsection{MICRONS CAVE cohorts}

MICRONS cohorts are queried from the CAVE \texttt{minnie65\_public} stack using
the \texttt{digital\_twin\_properties\_bcm\_coreg\_v4} table at materialization
version 1507. Three non-overlapping windows are used: discovery offset 0,
hold-out offset 1000, and hold-out offset 2000.

\subsection{MICRONS structure-function tests}

For each cohort, all directed non-self pairs are constructed. Connected pairs
are compared to non-connected pairs under random, distance-matched, and
degree-matched null models. The primary metric is readout-location similarity.
Tests are corrected by false discovery rate, and the manuscript endpoint is
accepted only if it passes in the discovery cohort and in both hold-out cohorts.

\subsection{Unit-cluster bootstrap}

Bootstrap intervals resample units rather than directed pairs. This decision
avoids treating the many pairwise comparisons derived from the same units as
fully independent observations. The bootstrap is therefore more conservative
than a pair-level resampling strategy.
