\section{Related work}
\label{sec:related-work}

MouseBrainBench lies at the intersection of neural-response prediction,
multiscale brain modelling, and functional connectomics. These areas share data
and biological targets, but they validate different propositions. This section
therefore reviews them according to the evidence they can provide rather than
as a single progression towards a complete digital brain.

\subsection{Standardized neural recordings and response prediction}
\label{subsec:recordings-prediction}

The Allen Brain Observatory has made standardized neurophysiological data from
the mouse visual system available at a scale that supports comparisons across
sessions, animals, and analysis pipelines \citep{siegle2021survey}. The Visual
\ac{vbn} resource is especially relevant because it combines
large-scale extracellular recordings with controlled visual and behavioural
conditions. Standardization and public access improve reproducibility, but the
recordings remain observations of activity. They do not provide the complete
synaptic graph of each recorded animal, and several latent circuits can generate
similar population responses.

Cross-resource integration also depends on anatomical standardization. The Allen
Mouse Brain \ac{ccf} provides a three-dimensional reference
atlas derived from 1675 brains at 10-$\mu$m voxel resolution
\citep{wang2020ccf}. This makes regional alignment and aggregation possible,
but a reference atlas is not an individual connectome. MouseBrainBench therefore
uses atlas identity as provenance and anatomical context rather than as evidence
that subject-specific wiring has been observed.

Whole-brain axonal tracing supplies a complementary mesoscale view. The Allen
Mouse Brain Connectivity Atlas established a brain-wide projection map
\citep{oh2014mesoscale}, and subsequent modelling inferred connectivity at
100-$\mu$m voxel resolution under spatial smoothness assumptions
\citep{knox2019connectome}. These resources are indispensable for regional and
voxel-scale models, but inferred projection strength is not a measured
single-neuron connectome. The distinction between measured and inferred
structure is therefore part of the provenance recorded by MouseBrainBench.

Sensorium introduced a complementary evaluation regime: predicting large-scale
neural responses to static natural images in mouse primary visual cortex
\citep{willeke2022sensorium}. Dynamic Sensorium extended the task to videos,
behavioural covariates, and out-of-distribution evaluation
\citep{turishcheva2023dynamic,turishcheva2024retrospective}. These competitions
have advanced model architectures, shared evaluation, and reproducible data
handling. Correlation with held-out neural responses is a meaningful measure of
functional approximation and can identify models that generalize better than
simple baselines.

Neural-response metrics must account for response variability and the
reliability ceiling of the recorded neurons \citep{schoppe2016performance}.
Recent foundation models extend prediction to new visual stimulus types and
provide functional descriptors for the MICRONS animal
\citep{wang2025foundation}. These advances strengthen predictive models as
experimental tools, while making it more important to state whether a reported
value measures raw prediction, reliability-normalized performance, or evidence
about circuit organization.

The term \emph{digital twin} is also used for neural response models that enable
experiments in stimulus space. Ustyuzhaninov et al. trained predictive models of
mouse \ac{v1} and used their outputs to characterize combinations of nonlinear
response properties \citep{ustyuzhaninov2022digitaltwin}. This approach shows
that a functional surrogate can support useful in silico experiments even when
it does not reproduce the underlying synaptic circuit. The distinction is
important because the scientific value of the surrogate does not require a
stronger mechanistic claim.

Predictive agreement at the level of individual neurons may also hide errors at
the population level. Liscai et al. found that visual-cortex digital twins can
match single-neuron responses while failing to reproduce relevant properties of
population-response geometry \citep{liscai2025population}. Their result provides
a direct example of metric-dependent validation. It supports the need to match
the evaluation target to the intended use and to avoid treating one predictive
score as evidence for every level of neural organization.

Nevertheless, predictive equivalence does not imply mechanistic equivalence.
Two architectures can generate similar held-out predictions while relying on
different internal representations. High response correlation therefore
supports a predictive-surrogate claim, conditional on reliability and test
design, but not by itself an anatomical or causal claim. MouseBrainBench retains
the established predictive metrics and adds an explicit boundary: predictive
results enter the audit as one evidence type and cannot satisfy absent topology
or directionality requirements.

\subsection{Multiscale simulation and digital-brain models}
\label{subsec:multiscale-simulation}

Biologically grounded simulators address a different problem. BioNet and \ac{bmtk}
provide interfaces and workflows for constructing networks from cellular and
connectivity information and executing them at different levels of resolution
\citep{gratiy2018bmtk,dai2020bmtk}. SONATA-compatible descriptions improve
portability between tools. At a coarser scale, \ac{tvb} models
population dynamics over connectomes and supports the study of large-scale
network behaviour \citep{tvb2013}. These systems demonstrate that simulation
engines, model formats, and numerical integration are already mature research
infrastructures.

Mouse-specific models further reduce the space in which a new simulator would
be novel. The Virtual Mouse Brain extends connectome-based population dynamics
to tracer- and imaging-derived mouse connectomes
\citep{melozzi2017virtualmouse}. At circuit scale, the Allen V1 model integrates
cellular, structural, and functional measurements into biophysically detailed
and point-neuron networks \citep{billeh2020v1}. MouseBrainBench does not compete
with either contribution. It evaluates whether outputs from such models support
the claim attached to them.

Digital Twin Brain proposals broaden this objective towards multiscale models
that connect biological and \acs{ai} \citep{xiong2023dtb}. The
MouseDTB submission is the closest whole-brain mouse proposal to the present
application domain \citep{chen2025mousedtb}. It constructs inferred
single-neuron connectivity from axonal projections and cell distributions, then
couples that structure to a large spiking network. Its evaluation includes \ac{bold}
signals in resting and olfactory-task conditions, parameter assimilation, and
behavioural responses. The work is currently reported as an ICLR 2026
submission. It should therefore be treated as an important contemporary system
rather than as an established benchmark.

MouseDTB and MouseBrainBench address different engineering questions. MouseDTB
asks whether a large inferred network can reproduce selected whole-brain and
behavioural observations. MouseBrainBench asks which claims are justified by
the available observations and controls. Inferred single-neuron connectivity
does not become an observed individual connectome solely because the model
matches \acs{bold} activity. Likewise, high \acs{bold} correlation does not establish that
the inferred wiring is uniquely identified. This difference prevents direct
competition while making MouseDTB a relevant candidate for future audit through
the same claim-aware protocol.

Their presence changes the novelty criterion for the present work. Rebuilding a
spiking engine or a connectome-based neural-mass simulator would duplicate
established technology without resolving whether the resulting output justifies
a biological claim. Conversely, successful numerical execution only establishes
that a model is computationally well posed under its assumptions. Biological
validity still depends on parameter provenance, target observability, null-model
comparisons, and evaluation under perturbations or independent data.

MouseBrainBench is therefore positioned above, rather than instead of, these
tools. An external simulator may produce candidate dynamics. The benchmark
records the target, controls, and evidence required to interpret them. This
separation also prevents the phrase \emph{digital twin} from being used as a
synonym for any executable brain model. The present framework reserves stronger
twin claims for systems that can demonstrate entity-specific calibration,
state-dependent prediction, and suitable interventional validation.

\subsection{Connectomics and structure-function analysis}
\label{subsec:connectomics}

Electron-microscopy connectomics offers the most direct local evidence about
synaptic wiring. \acs{microns} links a dense reconstruction of mouse visual cortex to
functional measurements and related cell properties \citep{bae2025microns}.
It is therefore suitable for asking whether connected neurons share functional
properties more strongly than comparable non-connected neurons. Unlike a
response-only dataset, \acs{microns} permits connectivity and functional descriptors
to be tested within a co-registered local population.

The combination of in vivo functional characterization and electron microscopy
predates \acs{microns}. Earlier work demonstrated that network anatomy and visual
physiology can be co-registered at smaller scale \citep{bock2011network}.
\acs{microns} changes the coverage of the analysis, but it does not remove the need to
distinguish exploratory pairwise associations from independently validated
connectivity rules.

Access to a changing reconstruction introduces an additional reproducibility
problem: cell identities and annotations can change as proofreading progresses.
\ac{cave} addresses this issue through versioned segmentation, materialized
annotation snapshots, and reproducible queries \citep{dorkenwald2025cave}. For
this reason, MouseBrainBench records the \acs{microns} datastack, table, and
materialization version rather than citing a live query alone.

This general biological question is not new. Ding et al. used the \acs{microns}
resource to demonstrate like-to-like connectivity within and across cortical
layers and visual areas, separating feature and receptive-field components and
using axon--dendrite proximity controls \citep{ding2025wiring}. Their analysis
is biologically richer than the compact endpoint used in this paper. Therefore,
the connected-pair result reported here must not be interpreted as the first
evidence for functional specificity of cortical connectivity.

This opportunity comes with important confounds. Connection probability and
functional similarity both vary with spatial separation. Neuronal degree,
cortical area, cell class, and sampling quality can also influence pairwise
comparisons. A connected-versus-random contrast is consequently insufficient.
Any defensible structure-function result must show that the effect survives
appropriate matched controls, multiplicity correction, and resampling at the
level of biological units rather than treating all derived pairs as independent.
It must also replicate outside the cohort used to define or discover the
endpoint.

Previous connectomic work establishes both the value of \acs{microns} and the broad
structure-function phenomenon. The present contribution is not the rediscovery
that structure and function are related. \acs{microns} is used as an external-validity
case: MouseBrainBench asks whether a simpler declared endpoint recovers a
compatible association under a fixed, reproducible audit and whether its
interpretation remains bounded. This is why the manuscript reports a replicated
local association, cites the stronger prior result, and explicitly withholds
biological-priority, causal, behavioural, and whole-brain conclusions.

The reviewed infrastructures provide data, simulators, or predictive
benchmarks, but they do not share a common decision layer for controlling the
scope of claims across these evidential regimes. Leaderboards prioritize
predictive performance. Simulators prioritize executable biological models.
Connectomic studies test local anatomical hypotheses. Comparing their headline
metrics directly would be scientifically incoherent.

MouseBrainBench fills this narrower gap by defining the claim before selecting
the evidence, preserving negative controls, and recording failure as an
informative outcome. Its novelty must therefore be judged as methodological:
whether the conjunctive audit distinguishes known failure modes and produces a
replicable, more tightly bounded conclusion than correlation-only evaluation.
The experiments in Section~\ref{sec:evaluation} are designed to test precisely
that claim.
