\section{Results}

The results follow the evidential hierarchy of the benchmark. Synthetic systems
test the decision logic under known truth. Allen tests whether the framework
rejects a mechanism despite highly reproducible observations. Sensorium tests
whether predictive gains are retained without being promoted to causal
evidence. MICRONS first exposes an aggregate failure and then supplies the
replicated local structure--function endpoint.

\subsection{MIS recovers synthetic truth and rejects incomplete mechanisms}

The known-truth experiment behaved as specified. The directed system passed all
three evidence blocks and obtained $M=1.0$. The common-drive system passed
reproducibility but failed topology specificity and directionality, yielding
$M=0.333$. The topology-only system passed reproducibility and topology but
failed directionality ($M=0.667$). The direction-only system contained temporal
information but failed topology specificity ($M=0.719$). Only the system
containing the complete designed mechanism passed the conjunctive decision.

The component statistics reflected the intended construction. In the
directed-truth system, median cross-session and split-half correlations were
0.988 and 0.994. Its prediction exceeded the median permuted prediction by
0.783 and the disconnected prediction by 0.916; every regional pair had a
resolvable latency. By contrast, the common-drive system reached cross-session
correlation 0.962 but its candidate prediction was 0.0009 worse than the
permutation median and 0.0032 worse than the disconnected control.

The topology-only case retained advantages of 0.157 over permutations and 0.081
over the disconnected model, yet its resolvable-latency fraction was zero. The
direction-only system demonstrated the complementary failure: the timing block
passed, but the topology block did not distinguish the candidate from its
controls. These cases show why the continuous MIS value cannot replace the
conjunctive decision. Partial scores diagnose different failure modes; they are
not a ranking of mechanistic truth.

\subsection{Allen responses are reproducible but do not identify the proposed
mechanism}

Allen produced the intended real negative control. Median cross-mouse
correlation was 0.891, median split-half correlation was 0.990, and 75\% of
instances exceeded their own 95th-percentile null. All three reproducibility
criteria passed, giving that block a score of 1.0.

The topology results led to the opposite conclusion. The Allen-informed model
performed 0.0054 worse than the median permutation and outperformed only 21\%
of permutations. The lower bound of the paired bootstrap interval was
$-0.0075$. Its advantage over a disconnected model was 0.0035, well below the
fixed 0.05 threshold. None of the four topology criteria passed and the block
score was 0.073.

Directed identifiability also failed. Median cross-mouse and split-half latency
agreement, the fraction above the latency null, the resolvable pair fraction,
the fraction above the lead--lag null, and the nonzero lead--lag fraction were
all zero under their operational definitions. The complete MIS was 0.358 and
the decision was
\texttt{reproducible\_target\_without\_mechanistic\_identifiability}.

This decision was not caused by marginal failure at one boundary. Multiplying
all Allen thresholds by 0.75, 1.00, or 1.25 produced the same block pattern:
reproducibility passed, whereas topology specificity and directed
identifiability failed. The result does not imply that the mouse visual system
lacks directed organization. It means that the selected population target and
model comparison do not identify that organization.

\subsection{Sensorium supports prediction without mechanistic passage}

Static Sensorium showed consistent predictive utility. Across seven validation
mice, median best correlation was 0.328, median improvement over mean response
was 0.102, and median improvement over the scrambled-stimulus control was
0.077. Reliability cannot be estimated from non-repeated validation stimuli and
was therefore not assigned a favorable value.

In the five repeated-test mice, median reliability was 0.642 and median best
predictive correlation was 0.341. Improvements over mean response and scrambled
stimulus were 0.095 and 0.068. The separate topographic constraint passed in all
five datasets, with median Spearman association 0.177 and median effect over its
null 0.178. This is partial structural evidence, but it does not identify a
synaptic graph or causal direction; no static Sensorium mouse passed the complete
MIS.

The sensitivity grid makes this boundary explicit. Reliability and topographic
evidence passed all six combinations using reliability thresholds 0.50 or 0.60
and topographic thresholds from 0.05 to 0.15. The partial-positive decision
failed at reliability threshold 0.70. Thus, topographic evidence was stable over
the tested range, whereas the strength assigned to reliability depends on a
threshold that brackets the observed median.

Dynamic Sensorium produced best per-mouse correlations from 0.427 to 0.526 in
the current public cohort. The temporal filterbank won one mouse, temporal
singular-value decomposition won two, and the compact multilayer perceptron won
two. The filterbank improved over mean response in all five mice, with median
gain 0.031.

The multilayer perceptron gain was more modest when compared with a serious
temporal baseline. Its median gain was 0.043 over mean response but 0.0039 over
temporal singular-value decomposition. In the legacy out-of-distribution
cohort, the corresponding gains were 0.036 and 0.0022. At a minimum effect of
0.01, the multilayer perceptron beat temporal singular-value decomposition in
only one of five mice in each cohort. Increasing complexity therefore improved
the simple baseline more consistently than it improved a transparent temporal
model.

The bounded official-architecture run completed through the official Sensorium
loader and model factory for five mice. Correlations ranged from 0.011 to 0.035
under the deliberately restricted training budget. The artifact establishes
software interoperability, not competitive model quality. The audit accordingly
marks it as a bounded integration control and blocks state-of-the-art or
leaderboard-equivalent interpretations.

\subsection{The aggregate MICRONS analysis prevents an initial overclaim}

The first expanded MICRONS analysis compared connected pairs with random,
distance-matched, and degree-matched non-connected pairs at aggregate level.
Connected pairs were more functionally similar than random pairs and
degree-matched controls, but the contrast did not survive distance matching.
Because cortical distance influences both connection probability and functional
organization, the aggregate result was classified as negative or inconclusive.

This failure altered the analysis rather than being discarded. It showed that a
random null was too permissive and that the aggregate association could not be
separated from spatial embedding. The expanded result was retained as a stress
test, and subsequent analyses were stratified by functional descriptor and
anatomical context. Readout-location similarity was fixed as the primary
endpoint for the replication package after this diagnostic stage.

Had only the random or degree-matched contrast been reported, the same data
could have supported an overstated positive conclusion. Requiring the
distance-matched gate changed the outcome and forced a narrower hypothesis. This
is a direct example of the claim-audit logic affecting scientific inference
rather than merely relabelling a result after analysis.

\subsection{MICRONS yields a replicated local structure--function association}

The three MICRONS cohorts jointly contain 2991 units, 6575 loaded synapses, and
5943 unique connected directed pairs. Discovery and the two hold-outs contain
2095, 1926, and 1922 connected pairs among 999000, 983072, and 997002 candidate
directed non-self pairs. The primary endpoint passed both distance- and
degree-matched tests in every cohort (Table~\ref{tab:microns-primary}).

\input{tables/microns_primary}

For distance-matched controls, connected-pair advantages in readout-location
similarity were 0.0215, 0.0210, and 0.0179, with one-sided
false-discovery-rate-adjusted $q$ values of 0.0076, 0.0070, and 0.0087.
Degree-matched effects were 0.0403, 0.0380, and 0.0286, with adjusted $q$
values of 0.0053, 0.0062, and 0.0083. The effect retained its sign and
approximate magnitude outside the discovery window.

The discovery family contained 174 statistical tests across functional
descriptors and strata, of which 28 were confirmed positive after
false-discovery-rate correction. The two hold-outs contained 162 and 174 tests,
with 30 and 25 confirmed positives. These counts describe the multiplicity
family and are not presented as independent discoveries. The manuscript claim
rests on the fixed readout-location endpoint and its replication. Bootstrap
stability is summarized in Table~\ref{tab:microns-bootstrap}.

\input{tables/microns_bootstrap}

Unit-cluster bootstrap intervals remained strictly positive. Distance-matched
lower bounds were 0.0166, 0.0142, and 0.0115; degree-matched lower bounds were
0.0349, 0.0321, and 0.0229. Resampling units rather than pairs reduces the false
precision that would result from treating pairwise observations sharing neurons
as independent. All 300 requested replicates were usable in each cohort.

The result is narrower than a causal mechanism. Readout location is a functional
descriptor associated with connectivity after two major controls, but unmeasured
cell type, layer, area, morphology, or developmental organization may
contribute. The analysis does not perturb edges or establish necessity.

Nor is the association a new biological wiring rule. Ding et al. reported a
substantially richer MICRONS analysis of like-to-like connectivity using
receptive-field, feature, and axon--dendrite geometry controls
\citep{ding2025wiring}. Agreement with that work provides an external-validity
case for MouseBrainBench. The potential novelty lies in applying one explicit
claim-control protocol across synthetic, Allen, Sensorium, and MICRONS regimes,
not in priority over the underlying biological phenomenon.

\subsection{The publication freeze preserves different evidential outcomes}

The final publication artifact admitted separation of predictive, reliability,
out-of-distribution, and mechanistic evidence; Allen as a reproducible negative
mechanistic case; Sensorium as predictive evidence; partial reliability and
topographic evidence in static Sensorium; bounded execution of the official
Sensorium stack; and a replicated local MICRONS readout-location association.

The same artifact blocked a complete mouse-brain digital twin, a
state-of-the-art Sensorium predictor, causal identifiability in Dynamic
Sensorium, a causal interpretation of the MICRONS association, and any
whole-brain extrapolation from the local volume. Replication made the package a
Q1 manuscript candidate but did not remove these restrictions.

The cross-regime result is therefore not that one model achieved the highest
score. It is that the same audit represented three outcomes without semantic
promotion: reproducible but non-identifiable, predictive but non-mechanistic,
and locally structure--function positive but observational. These distinctions
form the empirical support for the methodological contribution.
