\documentclass[a4paper,fleqn]{cas-sc}

\usepackage[authoryear]{natbib}
\usepackage{amsmath,amssymb,booktabs,tabularx}
\usepackage{url}

\journal{Engineering Applications of Artificial Intelligence}

\begin{document}
\let\WriteBookmarks\relax
\def\floatpagepagefraction{1}
\def\textpagefraction{.001}


\shorttitle{Claim-aware validation of mouse-brain digital models}
\shortauthors{Fern\'andez-Isabel}

\title[mode=title]{MouseBrainBench: Claim-aware verification and validation of
artificial intelligence models for partial digital representations of the
mouse brain}


\author[1]{Alberto Fern\'andez-Isabel}
\cormark[1]
\ead{alberto.fernandez.isabel@urjc.es}
\credit{Conceptualization, Methodology, Software, Validation, Formal analysis,
Investigation, Data curation, Writing--original draft, Writing--review and
editing, Project administration}
\affiliation[1]{organization={Data Science Laboratory, Rey Juan Carlos University},
                addressline={C/ Tulip\'an, s/n},
                city={M\'ostoles},
                postcode={28933},
                state={Madrid},
                country={Spain}}
\cortext[1]{Corresponding author}


\begin{abstract}
Artificial intelligence is increasingly used to construct predictive and
simulation-based models of complex scientific systems. However, model
evaluation is commonly reduced to predictive accuracy, even when the intended
claim concerns reproducibility, structural fidelity, or mechanism. This paper
introduces MouseBrainBench, an engineering framework for verification and
validation of partial digital models of the mouse brain. Its contribution to
artificial intelligence is a claim-aware benchmark architecture that represents
prediction, reproducibility, topology specificity, directed identifiability,
structure--function agreement, and computational cost as non-interchangeable
evidence blocks. A Mechanistic Identifiability Score provides descriptive
diagnostics, while a conjunctive decision rule prevents high performance in one
block from compensating for missing evidence in another.

The framework is evaluated with controlled simulations and three families of
public data. Synthetic systems confirm that the decision rule rejects common
drive, topology without temporal direction, and direction without
topology-specific prediction. Allen Visual Behavior Neuropixels provides a real
negative control: median cross-mouse and split-half correlations reach 0.891 and
0.990, but topology-specific and directed-identifiability criteria fail. Static
and Dynamic Sensorium demonstrate predictive gains without corresponding
mechanistic evidence. MICRONS provides an external structure--function
validation case. Across a discovery cohort and two non-overlapping hold-outs
containing 2991 co-registered units and 5943 connected directed pairs,
connected pairs exhibit greater readout-location similarity than distance- and
degree-matched controls. Distance-matched effects are 0.0215, 0.0210, and
0.0179, with positive unit-cluster bootstrap intervals in all cohorts.

MouseBrainBench does not claim a complete or causal digital brain. It provides
a reproducible engineering layer for deciding which claims about AI-based
scientific models are supported, partially supported, or blocked by the
available evidence.
\end{abstract}

\begin{highlights}
\item A claim-aware benchmark separates prediction from mechanistic evidence.
\item Conjunctive evidence gates prevent compensation across validation criteria.
\item Allen and Sensorium expose reproducible or predictive non-mechanistic cases.
\item A MICRONS endpoint replicates across two non-overlapping hold-out cohorts.
\item Decisions remain linked to versioned artifacts and explicitly blocked claims.
\end{highlights}

\begin{keywords}
artificial intelligence validation \sep scientific machine learning \sep
digital twins \sep computational neuroscience \sep benchmarking \sep
mechanistic identifiability
\end{keywords}


\maketitle


\section{Introduction}

The construction of computational representations of biological systems has
become an important meeting point between Artificial Intelligence and
neuroscience. In the laboratory mouse (\textit{Mus musculus}), standardized
electrophysiology, visual-response
benchmarks, anatomical atlases, and electron-microscopy reconstructions are now
available through public infrastructures. These resources make it possible to
study the same biological system at markedly different scales, from population
responses recorded across brain regions to synaptic connectivity within a
restricted cortical volume. They also make it technically possible to build
models that reproduce selected observations with increasing accuracy.

This progress creates a problem of interpretation. A model that predicts neural
responses is not necessarily a model that identifies the circuit producing
them. A response can be stable across animals while remaining compatible with
many alternative topologies. Likewise, a network can preserve an anatomical
graph without establishing that its directed dynamics are identifiable from the
chosen observations. Predictive performance, empirical reliability, anatomical
plausibility, and causal explanation therefore answer different questions. A
single correlation coefficient cannot adjudicate among them.

The distinction is particularly important for digital-twin terminology. A
digital twin normally implies correspondence with a particular physical entity,
continuous or repeated calibration, and the capacity to reproduce relevant
states under intervention. Current mouse-brain resources do not jointly provide
complete synaptic connectivity, full-brain dynamics, behaviour, and
interventional validation for an individual animal. MICRONS offers exceptional
synaptic and functional detail, but only for a local visual-cortical volume
\citep{bae2025microns}. Allen datasets provide broad and standardized neural
recordings, but not the individual synaptic graph that generated each response
\citep{siegle2021survey,deVries2022AllenSharing}. Consequently, referring to
either resource as a complete mouse-brain twin would exceed the evidence.

Existing tools solve other essential parts of the problem. BMTK provides
infrastructure for constructing and simulating multiscale neural networks
\citep{dai2020bmtk}; The Virtual Brain supports connectome-based population
dynamics \citep{tvb2013}; and Sensorium provides common predictive tasks for
mouse visual cortex \citep{willeke2022sensorium,turishcheva2023dynamic}. Their
existence means that another simulator or another visual prediction benchmark
would have limited novelty. What remains insufficiently formalized is the audit
between a reported metric and the biological claim attached to it: what must be
observed before reproducible prediction becomes evidence for a particular
mechanism?

MouseBrainBench addresses this question. It is not a new full-brain simulator
and it does not claim to instantiate a complete digital twin. It is a software
and experimental framework that assigns different public datasets to explicit
evidential roles, applies matched null models and hold-out replication, and
records which interpretations pass or fail. Its Mechanistic Identifiability
Score (MIS) is a conjunctive decision rule over non-interchangeable evidence
blocks. The numerical component summarizes support within a block, whereas the
final decision requires every block relevant to the proposed mechanism to pass.

The evaluation deliberately combines positive and negative outcomes. Four
synthetic systems test whether MIS recovers known truth and rejects common-drive
or incomplete mechanisms. Allen Visual Behavior Neuropixels tests whether the
framework can retain a strongly reproducible target while rejecting unsupported
topological and directional claims. Static and Dynamic Sensorium test modern
predictive models without promoting correlation to causation. Finally, MICRONS
tests a prespecified local structure-function endpoint against distance- and
degree-matched nulls in a discovery cohort and two non-overlapping hold-outs.
This sequence is intended to test the audit itself, not to assemble unrelated
dataset demonstrations.

The remainder of the paper follows this argument. Section~2 states the research
contributions. Section~3 analyses the relevant predictive, simulation, and
connectomic literature. Section~4 defines the framework and the MIS decision
rule. Section~5 describes the experimental protocol and evidential role of each
dataset. Section~6 presents and discusses the results, including negative
evidence and sensitivity limits. Section~7 concludes the work and delimits the
claims that remain outside the present study.

\section{Research contributions}

The main contribution is not a new mouse-brain simulator. It is a falsifiable
procedure for determining what kind of claim is supported by a partial digital
model. This contribution is developed through the following elements:

\begin{itemize}
  \item A claim-audit formulation that separates prediction, reproducibility,
  topology specificity, directed identifiability, local structure-function
  association, and computational cost. The output is not only a ranking: it is
  an explicit set of admitted and rejected interpretations.

  \item A formal MIS composed of non-interchangeable evidence blocks and fixed
  directional thresholds. The rule is tested with four known-truth systems that
  isolate common drive, topology without direction, and direction without
  topology specificity.

  \item A cross-resource protocol that assigns Allen, Sensorium, Dynamic
  Sensorium, and MICRONS distinct roles instead of treating heterogeneous data
  as interchangeable validation. This design provides a real negative control,
  predictive controls, and a local anatomical-functional test.

  \item An external-validity case in MICRONS obtained with a specified primary
  endpoint, distance- and degree-matched null models, false-discovery-rate
  correction, unit-cluster bootstrap intervals, and two non-overlapping hold-out
  cohorts. It tests reproducible recovery of a published broad phenomenon rather
  than claiming a new biological wiring rule.

  \item A reproducible publication freeze that binds numerical artifacts to
  allowed and blocked claims. In the current study it admits a replicated local
  observational structure-function association, while rejecting causal,
  behavioural, whole-brain, and state-of-the-art predictor claims.
\end{itemize}

These elements define the scope of the work. MouseBrainBench evaluates the
relationship between evidence and interpretation; it does not establish that a
single score is universally sufficient for mechanistic neuroscience. The MIS
thresholds and evidence blocks must be specified for the target claim before
evaluation, and failure means that the particular claim is unsupported, not
that the dataset or model is devoid of scientific value.

\section{Related work}

MouseBrainBench lies at the intersection of neural-response prediction,
multiscale brain modelling, and functional connectomics. These areas share data
and biological targets, but they validate different propositions. This section
therefore reviews them according to the evidence they can provide rather than
as a single progression towards a complete digital brain.

\subsection{Standardized neural recordings and response prediction}

The Allen Brain Observatory has made standardized neurophysiological data from
the mouse visual system available at a scale that supports comparisons across
sessions, animals, and analysis pipelines \citep{siegle2021survey}. The Visual
Behavior Neuropixels resource is especially relevant because it combines
large-scale extracellular recordings with controlled visual and behavioural
conditions. Standardization and public access improve reproducibility, but the
recordings remain observations of activity. They do not provide the complete
synaptic graph of each recorded animal, and several latent circuits can generate
similar population responses.

Cross-resource integration also depends on anatomical standardization. The Allen
Mouse Brain Common Coordinate Framework provides a three-dimensional reference
atlas derived from 1675 brains at 10-$\mu$m voxel resolution
\citep{wang2020ccf}. This makes regional alignment and aggregation possible,
but a reference atlas is not an individual connectome. MouseBrainBench therefore
uses atlas identity as provenance and anatomical context rather than as evidence
that subject-specific wiring has been observed.

Whole-brain axonal tracing supplies a complementary mesoscale view. The Allen
Mouse Brain Connectivity Atlas established a brain-wide projection map
\citep{oh2014mesoscale}, and subsequent modelling inferred connectivity at
100-$\mu$m voxel resolution under spatial smoothness assumptions
\citep{knox2019connectome}. These resources are indispensable for regional and
voxel-scale models, but inferred projection strength is not a measured
single-neuron connectome. The distinction between measured and inferred
structure is therefore part of the provenance recorded by MouseBrainBench.

Sensorium introduced a complementary evaluation regime: predicting large-scale
neural responses to static natural images in mouse primary visual cortex
\citep{willeke2022sensorium}. Dynamic Sensorium extended the task to videos,
behavioural covariates, and out-of-distribution evaluation
\citep{turishcheva2023dynamic,turishcheva2024retrospective}. These competitions
have advanced model architectures, shared evaluation, and reproducible data
handling. Correlation with held-out neural responses is a meaningful measure of
functional approximation and can identify models that generalize better than
simple baselines.

Neural-response metrics must account for response variability and the
reliability ceiling of the recorded neurons \citep{schoppe2016performance}.
Recent foundation models extend prediction to new visual stimulus types and
provide functional descriptors for the MICRONS animal
\citep{wang2025foundation}. These advances strengthen predictive models as
experimental tools, while making it more important to state whether a reported
value measures raw prediction, reliability-normalized performance, or evidence
about circuit organization.

Nevertheless, predictive equivalence does not imply mechanistic equivalence.
Two architectures can generate similar held-out predictions while relying on
different internal representations. High response correlation therefore
supports a predictive-surrogate claim, conditional on reliability and test
design, but not by itself an anatomical or causal claim. MouseBrainBench retains
the established predictive metrics and adds an explicit boundary: predictive
results enter the audit as one evidence type and cannot satisfy absent topology
or directionality requirements.

\subsection{Multiscale simulation and digital-brain models}

Biologically grounded simulators address a different problem. BioNet and BMTK
provide interfaces and workflows for constructing networks from cellular and
connectivity information and executing them at different levels of resolution
\citep{gratiy2018bmtk,dai2020bmtk}. SONATA-compatible descriptions improve
portability between tools. At a coarser scale, The Virtual Brain models
population dynamics over connectomes and supports the study of large-scale
network behaviour \citep{tvb2013}. These systems demonstrate that simulation
engines, model formats, and numerical integration are already mature research
infrastructures.

Mouse-specific models further reduce the space in which a new simulator would
be novel. The Virtual Mouse Brain extends connectome-based population dynamics
to tracer- and imaging-derived mouse connectomes
\citep{melozzi2017virtualmouse}. At circuit scale, the Allen V1 model integrates
cellular, structural, and functional measurements into biophysically detailed
and point-neuron networks \citep{billeh2020v1}. MouseBrainBench does not compete
with either contribution; it evaluates whether outputs from such models support
the claim attached to them.

Their presence changes the novelty criterion for the present work. Rebuilding a
spiking engine or a connectome-based neural-mass simulator would duplicate
established technology without resolving whether the resulting output justifies
a biological claim. Conversely, successful numerical execution only establishes
that a model is computationally well posed under its assumptions. Biological
validity still depends on parameter provenance, target observability, null-model
comparisons, and evaluation under perturbations or independent data.

MouseBrainBench is therefore positioned above, rather than instead of, these
tools. An external simulator may produce candidate dynamics; the benchmark
records the target, controls, and evidence required to interpret them. This
separation also prevents the phrase \emph{digital twin} from being used as a
synonym for any executable brain model. The present framework reserves stronger
twin claims for systems that can demonstrate entity-specific calibration,
state-dependent prediction, and suitable interventional validation.

\subsection{Connectomics and structure-function analysis}

Electron-microscopy connectomics offers the most direct local evidence about
synaptic wiring. MICRONS links a dense reconstruction of mouse visual cortex to
functional measurements and related cell properties \citep{bae2025microns}.
It is therefore suitable for asking whether connected neurons share functional
properties more strongly than comparable non-connected neurons. Unlike a
response-only dataset, MICRONS permits connectivity and functional descriptors
to be tested within a co-registered local population.

The combination of in vivo functional characterization and electron microscopy
predates MICRONS. Earlier work demonstrated that network anatomy and visual
physiology can be co-registered at smaller scale \citep{bock2011network}.
MICRONS changes the coverage of the analysis, but it does not remove the need to
distinguish exploratory pairwise associations from independently validated
connectivity rules.

Access to a changing reconstruction introduces an additional reproducibility
problem: cell identities and annotations can change as proofreading progresses.
CAVE addresses this issue through versioned segmentation, materialized
annotation snapshots, and reproducible queries \citep{dorkenwald2025cave}. For
this reason, MouseBrainBench records the MICRONS datastack, table, and
materialization version rather than citing a live query alone.

This general biological question is not new. Ding et al. used the MICRONS
resource to demonstrate like-to-like connectivity within and across cortical
layers and visual areas, separating feature and receptive-field components and
using axon--dendrite proximity controls \citep{ding2025wiring}. Their analysis
is biologically richer than the compact endpoint used in this paper. Therefore,
the connected-pair result reported here must not be interpreted as the first
evidence for functional specificity of cortical connectivity.

This opportunity comes with important confounds. Connection probability and
functional similarity both vary with spatial separation; neuronal degree,
cortical area, cell class, and sampling quality can also influence pairwise
comparisons. A connected-versus-random contrast is consequently insufficient.
Any defensible structure-function result must show that the effect survives
appropriate matched controls, multiplicity correction, and resampling at the
level of biological units rather than treating all derived pairs as independent.
It must also replicate outside the cohort used to define or discover the
endpoint.

Previous connectomic work establishes both the value of MICRONS and the broad
structure-function phenomenon. The present contribution is not the rediscovery
that structure and function are related. MICRONS is used as an external-validity
case: MouseBrainBench asks whether a simpler declared endpoint recovers a
compatible association under a fixed, reproducible audit and whether its
interpretation remains bounded. This is why the manuscript reports a replicated
local association, cites the stronger prior result, and explicitly withholds
biological-priority, causal, behavioural, and whole-brain conclusions.

\subsection{Gap addressed by MouseBrainBench}

The reviewed infrastructures provide data, simulators, or predictive
benchmarks, but they do not share a common decision layer for controlling the
scope of claims across these evidential regimes. Leaderboards prioritize
predictive performance; simulators prioritize executable biological models; and
connectomic studies test local anatomical hypotheses. Comparing their headline
metrics directly would be scientifically incoherent.

MouseBrainBench fills this narrower gap by defining the claim before selecting
the evidence, preserving negative controls, and recording failure as an
informative outcome. Its novelty must therefore be judged as methodological:
whether the conjunctive audit distinguishes known failure modes and produces a
replicable, more tightly bounded conclusion than correlation-only evaluation.
The experiments in Sections~5 and~6 are designed to test precisely that claim.

\section{Proposed framework}

MouseBrainBench is a claim-aware verification and validation framework for
partial digital models of the mouse brain. It is not a neuronal simulator and
does not replace domain infrastructures such as BMTK, SONATA, The Virtual
Brain, Sensorium, or CAVE. These systems supply models, predictions, or data.
MouseBrainBench supplies the engineering contract that links their outputs to a
bounded scientific claim.

\subsection{Problem formulation}

Let $\mathcal{M}$ denote a computational model, $\mathcal{D}$ an evaluation
dataset, and $C$ a claim made about their relationship. Conventional
benchmarking usually estimates a predictive utility
$U(\mathcal{M},\mathcal{D})$ and ranks models using that scalar. This is
appropriate when $C$ is limited to held-out prediction under the same
evaluation protocol. It is insufficient when $C$ states that $\mathcal{M}$
recovers anatomical organization, directed dynamics, or a biological mechanism.
Several models may be predictively equivalent while encoding incompatible
internal structures.

MouseBrainBench represents a benchmark task as
\[
\mathcal{T} =
\left(C,\mathcal{O},\mathcal{B},\mathcal{N},\mathcal{R},\mathcal{P}\right),
\]
where $\mathcal{O}$ is the set of observable quantities, $\mathcal{B}$ contains
the evidence blocks required by the claim, $\mathcal{N}$ defines null and
alternative models, $\mathcal{R}$ contains decision rules, and $\mathcal{P}$
records provenance. The task is valid only when every quantity needed by
$\mathcal{B}$ is observable in $\mathcal{D}$. Missing evidence is not imputed
from predictive performance.

This formulation makes the claim an input to evaluation rather than a
post-hoc interpretation. A Sensorium predictor can be accepted as a functional
surrogate while being rejected as an anatomical explanation. A local MICRONS
association can pass structure--function controls while remaining insufficient
for a causal or whole-brain claim. The same numerical result can therefore have
different admissible interpretations depending on $C$.

\subsection{Evidence blocks and non-compensatory decision}

An evidence block contains criteria that answer one validation question. The
canonical configuration uses reproducibility, topology specificity, and
directed identifiability. Predictive tasks additionally report held-out
accuracy, out-of-distribution performance, and empirical reliability.
Structure--function tasks introduce matched anatomical controls and independent
replication. Computational cost is recorded as an engineering property but does
not substitute for biological evidence.

For criterion $j$ in block $i$, let $x_{ij}$ be the observed value,
$t_{ij}$ its fixed threshold, and $d_{ij}$ the comparison direction. Criterion
passage is
\[
p_{ij}=\mathbb{I}\!\left[x_{ij}\ d_{ij}\ t_{ij}\right].
\]
For a criterion requiring a value larger than its threshold, the normalized
diagnostic score is
\[
s_{ij}=\min\!\left(1,\max\!\left(0,\frac{x_{ij}}{t_{ij}}\right)\right).
\]
The inverse ratio is used when smaller values provide stronger evidence.
Zero-threshold criteria receive one when they pass and zero otherwise.

The descriptive score of evidence block $i$ is
\[
S_i=\frac{1}{n_i}\sum_{j=1}^{n_i}s_{ij},
\]
whereas block passage is $P_i=\bigwedge_j p_{ij}$. For $m$ blocks, the reported
Mechanistic Identifiability Score (MIS) is
\[
M=\frac{1}{m}\sum_{i=1}^{m}S_i,
\]
and the claim decision is
\[
P_{\mathrm{MIS}}=\bigwedge_{i=1}^{m}P_i.
\]

The separation between $M$ and $P_{\mathrm{MIS}}$ is intentional. $M$ helps
locate weak blocks and compare sensitivity scenarios, but it is not an additive
currency. Reproducibility cannot compensate for a topology that performs like a
permuted graph. Directed timing cannot compensate for failure to distinguish
the proposed regional organization. This non-compensatory behavior differentiates
the framework from a weighted average in which unrelated evidence can cancel
out.

\subsection{Claim classes}

MouseBrainBench distinguishes four claim classes used in this study. A
\emph{predictive claim} states that a model improves held-out neural-response
prediction over specified baselines. A \emph{reproducibility claim} states that
the target or effect is stable across repeated stimuli, sessions, animals, or
partitions. A \emph{structure--function claim} states that a structural relation
is associated with a functional descriptor after declared anatomical controls.
A \emph{mechanistic-identifiability claim} additionally requires that the
proposed topology and direction can be distinguished from plausible
alternatives.

These classes form a partial order rather than a linear quality scale.
Mechanistic identifiability requires reproducible observations, but useful
prediction does not automatically imply structural fidelity. A
structure--function association can be stronger than an uncontrolled
correlation and still weaker than intervention-supported causality. The audit
preserves these intermediate outcomes instead of forcing every benchmark into a
binary useful/useless label.

The framework also defines blocked claim classes. Whole-brain digital-twin
claims require evidence across brain-wide structure, dynamics, calibration, and
perturbation that none of the local resources used here jointly provide.
Behavioural-twin claims require closed-loop or task-level validation.
Consciousness claims have no operational target in the present system. Blocking
these statements is a validation result, not a software limitation disguised as
future functionality.

\subsection{Software architecture}

The implementation follows a layered architecture. The data layer contains
source-specific loaders, typed schemas, cache policies, and provenance records.
The representation layer maps regions, neurons, functional units, directed
edges, stimuli, and responses into stable Python objects. The model layer
contains transparent dynamics and adapters for external predictors. The
validation layer implements reliability, anatomical, functional, perturbation,
cost, and MIS metrics. The orchestration layer executes benchmark plans and
freezes their claim status.

Adapters isolate external ecosystems from the benchmark core. Allen exporters
convert sessions into normalized analysis artifacts. Sensorium adapters expose
static and dynamic predictors through a common response contract. CAVE adapters
query materialized MICRONS tables and preserve datastack and materialization
identifiers. BMTK and TVB adapters are reserved for external simulation outputs;
the framework does not reimplement their numerical engines.

Each benchmark command writes a machine-readable JSON artifact and a
human-readable Markdown summary. Artifacts contain the software version, Git
revision, configuration, source identifiers, cohort size, random seed, metrics,
thresholds, block decisions, and final interpretation. A result can therefore
be traced from a manuscript statement to the exact data and code state that
generated it.

\subsection{Null models and controls}

Null models are part of the task definition. Region permutations test whether a
candidate organization contributes more than label assignment. Disconnected
models test whether interactions add information beyond independent regional
dynamics. Scrambled-stimulus controls break stimulus--response correspondence
while preserving marginal statistics. Distance-matched MICRONS controls account
for spatial embedding, and degree-matched controls account for unequal
representation of highly connected neurons.

Controls are selected according to the confound addressed by the claim. Passing
a weak random null does not imply passage of a stronger matched null. This
principle is visible in the MICRONS experiments: an aggregate effect passes
random and degree controls but fails distance matching, forcing the claim to
remain negative until a stratified endpoint replicates under both controls.

\subsection{Publication freeze and traceability}

The final engineering component is the publication freeze. It reads sealed
benchmark artifacts and produces explicit lists of admitted, partially
supported, and blocked claims. It does not recompute results or permit manual
replacement of headline metrics. Manuscript assertions are generated from this
contract, reducing the risk that exploratory findings are promoted during
writing.

The freeze also reports incomplete baselines and external dependencies. For
example, successful execution of a bounded official Sensorium model is recorded,
but the state-of-the-art claim remains blocked because the published training
budget and leaderboard protocol were not reproduced. This distinction allows
engineering integration to be documented without treating it as empirical
equivalence.

\subsection{Evidential role of each dataset}

No available resource contains every observation required for a complete
digital-twin claim. Synthetic systems test decision logic under known truth.
Allen provides standardized, multi-animal neural activity and a negative
mechanistic control. Static and Dynamic Sensorium provide modern predictive
tasks. MICRONS provides local co-registered structure and function. Their roles
are complementary rather than interchangeable.

Table~\ref{tab:dataset-roles} summarizes this design. A negative MIS result for
Allen or Sensorium does not imply low dataset quality. It states that the
particular observations used here do not identify the stronger mechanism. The
MICRONS result is positive only for a local observational association and does
not upgrade the other datasets or the complete system to a digital twin.

\begin{table}[t]
\centering
\caption{Dataset roles in MouseBrainBench. Each dataset contributes a distinct evidence layer rather than a generic positive claim.}
\label{tab:dataset-roles}
\small
\begin{tabularx}{\textwidth}{l l X X}
\toprule
Dataset & Scale & Role & Claim status \\
\midrule
Synthetic MIS & 4 controlled cases & Score sanity check & Method validation \\
Allen VBN & 20--21 sessions & Negative mechanistic control & Reproducible, not identifiable \\
Sensorium static & 7 validation mice & Predictive/reliability case & Predictive, not causal \\
Dynamic Sensorium & 5 current + 5 OOD mice & Temporal predictive stress test & Predictive, underpowered \\
MICRONS CAVE & 2991 units, 6575 synapses & Main empirical case & Replicated local association \\
\bottomrule
\end{tabularx}
\end{table}

\section{Experimental methodology}

\subsection{Study design and claim hierarchy}

The study was designed as a methodological validation rather than as a
competition between datasets or model architectures. Each dataset was assigned
a role before analysis according to the observations it contains. Synthetic
systems provide known truth; Allen Visual Behavior Neuropixels provides repeated
population activity without individual synaptic connectivity; Sensorium
provides standardized predictive targets; and MICRONS provides co-registered
local anatomy and functional descriptors. These resources are therefore not
treated as interchangeable samples from one population.

A benchmark instance contains five elements: a bounded claim, observable
quantities, required evidence blocks, decision criteria, and provenance. The
claim is formulated before metric selection. For example, the claim that a
specific directed topology is identifiable requires reproducibility, superiority
over plausible alternative topologies, and a resolvable directional signature.
By contrast, the claim that connectivity is locally associated with a functional
descriptor requires co-registered structure and function, controls for major
anatomical confounds, multiplicity correction, and independent replication. It
does not require temporal direction because it is explicitly observational.

The analysis follows a sequence of increasingly realistic tests. First, the
decision rule is evaluated under known truth. Second, a real negative control
tests whether reproducibility is incorrectly promoted to mechanism. Third,
predictive benchmarks test whether model accuracy is kept separate from
anatomical evidence. Finally, the MICRONS analysis tests whether the framework
can admit a positive but bounded structure--function claim. The publication
freeze is generated only after all component artifacts are available.

\subsection{Mechanistic Identifiability Score}

Let $C$ be a scientific claim and let $B(C)=\{B_1,\ldots,B_m\}$ be its
required evidence blocks. Each block $B_i$ contains criteria
$c_{ij}=(x_{ij},t_{ij},d_{ij})$, where $x_{ij}$ is an observed statistic,
$t_{ij}$ is a fixed threshold, and $d_{ij}$ specifies whether the statistic
must be greater than, greater than or equal to, less than, or less than or equal
to that threshold. Criterion passage is
\[
p_{ij}=\mathbb{I}\left[x_{ij}\ d_{ij}\ t_{ij}\right].
\]
For criteria requiring a larger value, the descriptive normalized score is
$s_{ij}=\min(1,\max(0,x_{ij}/t_{ij}))$. For criteria requiring a smaller
value, the inverse ratio is used. Zero-threshold criteria receive one when they
pass and zero otherwise.

The score of block $i$ is the arithmetic mean
\[
S_i=\frac{1}{n_i}\sum_{j=1}^{n_i}s_{ij},
\]
and the displayed MIS value is $M=m^{-1}\sum_i S_i$. However, a block passes
only when all of its criteria pass, $P_i=\bigwedge_j p_{ij}$, and the
mechanistic decision is
\[
P_{\mathrm{MIS}}=\bigwedge_{i=1}^{m}P_i.
\]
The continuous value $M$ is diagnostic; it does not override the conjunctive
decision. This prevents strong reproducibility from compensating for absent
topology specificity, or strong prediction from compensating for missing
directional evidence.

The canonical implementation contains three blocks. Reproducibility assesses
whether the empirical target is stable across animals, sessions, or independent
partitions. Topology specificity assesses whether the proposed structure
outperforms disconnected, permuted, transposed, or otherwise matched
alternatives. Directed identifiability assesses whether latency, lead--lag
organization, or an equivalent directional signature is resolved reliably. A
dataset-specific audit can add observational structure--function criteria, but
those criteria are not relabelled as causal directionality.

\subsection{Synthetic known-truth systems}

Each synthetic case contains 24 sessions, six regions, and 24 time samples. A
regional template is generated from Gaussian temporal profiles. Directed
latency is introduced by shifting the profile centre by 0.08 time units per
region. Region-specific structure is introduced through systematic changes in
amplitude and width. Session noise has standard deviation 0.08; independent
odd/even perturbations used for split-half evaluation have standard deviation
0.04. Fixed random seeds 11, 23, 31, and 43 identify the four cases.

Reproducibility is measured as the median correlation between individual
sessions and their mean template, median odd--even correlation, and the fraction
of sessions exceeding the 95th percentile of 200 within-session region
permutations. Topology specificity compares the candidate prediction with 100
region-permuted predictions and a disconnected template. Directionality is
measured as the fraction of regional pairs with distinct peak latencies and the
equivalent nonzero lead--lag fraction.

The four systems isolate complementary conditions. The directed-truth case
contains all required information. The common-drive case is reproducible but has
neither region-specific topology nor direction. The topology-without-direction
case contains structured regional amplitudes but simultaneous temporal peaks.
The direction-without-topology case contains staggered peaks while its candidate
prediction is intentionally no more topology-specific than the disconnected
alternative. Before execution, the expected decision was that only the
directed-truth case should pass.

\subsection{Allen Visual Behavior Neuropixels negative control}

Allen analyses use exported population-response artifacts from Visual Behavior
Neuropixels sessions. The reproducibility phase contains 20 successful sessions.
The anatomical test contains 20 sessions from 20 unique mice, while the
directional test contains 21 sessions from 19 mice. Session inclusion, regional
aggregation, time windows, and preprocessing are fixed in the corresponding
configuration and exported with the result artifacts.

The reproducibility block requires median cross-mouse correlation greater than
0.50, median split-half correlation greater than 0.70, and at least half of
sessions above their own 95th-percentile null. The topology block requires an
Allen-informed advantage greater than 0.05 over both the median permutation and
the disconnected model, at least 95\% of permutations outperformed, and a
positive lower bound for the paired bootstrap interval. These thresholds ask
whether anatomical organization contributes information beyond generic regional
dynamics.

The directional block uses six criteria. Median cross-mouse latency Kendall
correlation must exceed 0.30, median split-half latency correlation must exceed
0.50, and at least half the sessions must exceed their own latency null. At least
half the regional pairs must have resolvable latency, at least half the sessions
must exceed the lead--lag null, and the median nonzero lead--lag pair fraction
must be at least 0.25. Failure of this block does not imply absence of biological
directionality; it means that the selected target does not resolve it under the
specified procedure.

\subsection{Static Sensorium evaluation}

Static Sensorium evaluation uses seven validation mice and five mice with
repeated test stimuli. The response of each neuron is compared with transparent
predictors: a stimulus-independent mean response, ridge regression over stimulus
features, and ridge regression over stimulus and behavioural context. A
scrambled-stimulus control preserves response statistics while breaking the
stimulus--response assignment. Pearson correlation is computed on held-out
responses and summarized at the mouse level.

Repeated stimuli provide an empirical reliability estimate. Reliability is not
imputed for tiers without repetitions; those tiers remain informative for
prediction but cannot pass a repeat-reliability gate. The topographic constraint
tests whether similarity in response descriptors varies systematically with
recorded spatial organization. Its Spearman statistic and effect over a matched
null are reported separately from prediction because a topographic association
does not establish synaptic connectivity or causal direction.

The static benchmark is deliberately not scored as a binary model competition.
It asks whether stimulus-aware models beat mean and scrambled controls, whether
the response is repeatable where repetitions exist, and whether partial
topographic evidence is present. Mechanistic passage additionally requires
evidence unavailable in this configuration and is therefore expected to remain
negative.

\subsection{Dynamic Sensorium predictive stress test}

Dynamic Sensorium uses five current public mice and five legacy
out-of-distribution mice. Seven model families are represented in the comparator:
mean response, summary features, temporal filterbank, temporal singular-value
decomposition, random features, a compact PyTorch multilayer perceptron, and a
bounded model instantiated through the official Sensorium loader and model
factory. Predictions are paired by mouse to avoid interpreting between-animal
differences as model effects.

The transparent temporal models provide intermediate complexity between a mean
response and a neural network. The filterbank represents finite temporal
kernels. Temporal singular-value decomposition represents a low-rank temporal
subspace. The random-feature model tests whether nonlinear expansion alone
explains gains, and the multilayer perceptron tests a compact learned nonlinear
mapping. For each mouse, the comparator reports correlation, the winning model,
and paired differences between model families.

The official architecture was trained under a bounded local budget on Apple
Metal Performance Shaders. It verifies that the official software stack,
dataloader, model factory, training, and evaluation can execute in the
MouseBrainBench environment. It does not reproduce the published training
budget or leaderboard protocol and is therefore excluded from state-of-the-art
claims. This distinction is encoded in the audit rather than left as a narrative
qualification.

\subsection{MICRONS data acquisition and cohort construction}

MICRONS data are queried from the public \texttt{minnie65\_public} CAVE stack.
Functional descriptors come from
\texttt{digital\_twin\_properties\_bcm\_coreg\_v4}; structural edges are
queried at materialization version 1507. The acquisition artifact stores the
datastack, table, materialization, query limits, offsets, source paths, and
result counts. CAVE materializations provide a time-locked annotation state
despite continued proofreading \citep{dorkenwald2025cave}. A CAVE token and
acceptance of the service terms are required to reproduce remote queries.

Three non-overlapping ordered windows define discovery and replication. Offset
0 provides 1000 usable units, offset 1000 provides 992, and offset 2000 provides
999. The windows were not pooled before evaluation. Structural queries yielded
2267, 2161, and 2147 synapses, corresponding to 2095, 1926, and 1922 unique
connected directed pairs. All possible directed non-self pairs form the
candidate universe in each cohort.

The primary endpoint is the similarity of the digital-twin readout-location
descriptor for connected pairs. This endpoint was selected after the expanded
aggregate analysis failed its distance-matched gate and was then fixed for the
replication package. The discovery and hold-out artifacts are retained
separately. Additional functional descriptors and anatomical strata are treated
as secondary analyses and are not promoted to the manuscript's primary claim.

\subsection{Matched controls and statistical inference}

For each connected directed pair, three comparator families are constructed.
Random non-connected pairs estimate the unconditioned contrast. Distance-matched
non-connected pairs control for spatial separation because connection
probability and functional organization both vary with distance. Degree-matched
non-connected pairs control for unequal representation of neurons with many
incoming or outgoing edges. The reported effect, $\Delta$, is the connected
pair similarity minus the corresponding control similarity.

Tests are one-sided for the prespecified hypothesis that connected pairs are
more similar than non-connected controls. The Benjamini--Hochberg procedure
controls the false discovery rate over the family of tests within each
stratified analysis. The primary endpoint is considered confirmed only if both
distance- and degree-matched effects are positive after correction in discovery
and in each hold-out cohort.

Pairwise observations are dependent because one neuron contributes to many
pairs. Confidence intervals therefore use a unit-cluster bootstrap rather than
a naive pair bootstrap. Each of 300 samples resamples units with replacement,
reconstructs the eligible pair set, and recomputes matched effects. Seed 71 is
fixed in the reproducibility package. The 2.5th and 97.5th percentiles of usable replicates
form the 95\% interval.

\subsection{Sensitivity, provenance, and publication freeze}

Threshold sensitivity is evaluated without reselecting outcomes. Allen MIS
criteria are multiplied by 0.75, 1.00, and 1.25. Static Sensorium reliability
thresholds of 0.50, 0.60, and 0.70 are crossed with topographic thresholds of
0.05, 0.10, and 0.15. Dynamic Sensorium paired gains are examined at minimum
effects of 0, 0.002, 0.005, and 0.01. These checks determine whether the
qualitative interpretation depends on a single boundary.

The Python package separates loaders, typed schemas, connectivity, dynamics,
simulation, validation, baseline adapters, and visualization. Benchmark commands
write JSON and Markdown summaries containing software version, Git revision,
inputs, sample counts, criteria, and decisions. Automated tests cover criterion
direction, normalization, conjunctive block semantics, known-truth cases,
dataset adapters, and publication gates.

The publication-freeze command reads sealed artifacts and emits two explicit
lists: claims supported by the current evidence and claims that remain blocked.
It does not recalculate primary statistics or permit manual substitution of
headline values. This design keeps manuscript language synchronized with the
reproducible evidence package and makes any future change in claim status
auditable.

\subsection{Ethics and use of public animal data}

This study performs secondary computational analyses of publicly released data
and introduces no new animal procedures. Animal acquisition, experimental
protocols, welfare oversight, and institutional approvals are described in the
original Allen, Sensorium, Dynamic Sensorium, and MICRONS publications cited in
this manuscript. MouseBrainBench operates on released recordings, metadata, and
derived structural resources under their applicable access conditions. No
personally identifiable human data are used.

\section{Results}

The results follow the evidential hierarchy of the benchmark. Synthetic systems
test the decision logic under known truth. Allen tests whether the framework
rejects a mechanism despite highly reproducible observations. Sensorium tests
whether predictive gains are retained without being promoted to causal
evidence. MICRONS first exposes an aggregate failure and then supplies the
replicated local structure--function endpoint.

\subsection{MIS recovers synthetic truth and rejects incomplete mechanisms}

The known-truth experiment behaved as specified. The directed system passed all
three evidence blocks and obtained $M=1.0$. The common-drive system passed
reproducibility but failed topology specificity and directionality, yielding
$M=0.333$. The topology-only system passed reproducibility and topology but
failed directionality ($M=0.667$). The direction-only system contained temporal
information but failed topology specificity ($M=0.719$). Only the system
containing the complete designed mechanism passed the conjunctive decision.

The component statistics reflected the intended construction. In the
directed-truth system, median cross-session and split-half correlations were
0.988 and 0.994. Its prediction exceeded the median permuted prediction by
0.783 and the disconnected prediction by 0.916; every regional pair had a
resolvable latency. By contrast, the common-drive system reached cross-session
correlation 0.962 but its candidate prediction was 0.0009 worse than the
permutation median and 0.0032 worse than the disconnected control.

The topology-only case retained advantages of 0.157 over permutations and 0.081
over the disconnected model, yet its resolvable-latency fraction was zero. The
direction-only system demonstrated the complementary failure: the timing block
passed, but the topology block did not distinguish the candidate from its
controls. These cases show why the continuous MIS value cannot replace the
conjunctive decision. Partial scores diagnose different failure modes; they are
not a ranking of mechanistic truth.

\subsection{Allen responses are reproducible but do not identify the proposed
mechanism}

Allen produced the intended real negative control. Median cross-mouse
correlation was 0.891, median split-half correlation was 0.990, and 75\% of
instances exceeded their own 95th-percentile null. All three reproducibility
criteria passed, giving that block a score of 1.0.

The topology results led to the opposite conclusion. The Allen-informed model
performed 0.0054 worse than the median permutation and outperformed only 21\%
of permutations. The lower bound of the paired bootstrap interval was
$-0.0075$. Its advantage over a disconnected model was 0.0035, well below the
fixed 0.05 threshold. None of the four topology criteria passed and the block
score was 0.073.

Directed identifiability also failed. Median cross-mouse and split-half latency
agreement, the fraction above the latency null, the resolvable pair fraction,
the fraction above the lead--lag null, and the nonzero lead--lag fraction were
all zero under their operational definitions. The complete MIS was 0.358 and
the decision was
\texttt{reproducible\_target\_without\_mechanistic\_identifiability}.

This decision was not caused by marginal failure at one boundary. Multiplying
all Allen thresholds by 0.75, 1.00, or 1.25 produced the same block pattern:
reproducibility passed, whereas topology specificity and directed
identifiability failed. The result does not imply that the mouse visual system
lacks directed organization. It means that the selected population target and
model comparison do not identify that organization.

\subsection{Sensorium supports prediction without mechanistic passage}

Static Sensorium showed consistent predictive utility. Across seven validation
mice, median best correlation was 0.328, median improvement over mean response
was 0.102, and median improvement over the scrambled-stimulus control was
0.077. Reliability cannot be estimated from non-repeated validation stimuli and
was therefore not assigned a favorable value.

In the five repeated-test mice, median reliability was 0.642 and median best
predictive correlation was 0.341. Improvements over mean response and scrambled
stimulus were 0.095 and 0.068. The separate topographic constraint passed in all
five datasets, with median Spearman association 0.177 and median effect over its
null 0.178. This is partial structural evidence, but it does not identify a
synaptic graph or causal direction; no static Sensorium mouse passed the complete
MIS.

The sensitivity grid makes this boundary explicit. Reliability and topographic
evidence passed all six combinations using reliability thresholds 0.50 or 0.60
and topographic thresholds from 0.05 to 0.15. The partial-positive decision
failed at reliability threshold 0.70. Thus, topographic evidence was stable over
the tested range, whereas the strength assigned to reliability depends on a
threshold that brackets the observed median.

Dynamic Sensorium produced best per-mouse correlations from 0.427 to 0.526 in
the current public cohort. The temporal filterbank won one mouse, temporal
singular-value decomposition won two, and the compact multilayer perceptron won
two. The filterbank improved over mean response in all five mice, with median
gain 0.031.

The multilayer perceptron gain was more modest when compared with a serious
temporal baseline. Its median gain was 0.043 over mean response but 0.0039 over
temporal singular-value decomposition. In the legacy out-of-distribution
cohort, the corresponding gains were 0.036 and 0.0022. At a minimum effect of
0.01, the multilayer perceptron beat temporal singular-value decomposition in
only one of five mice in each cohort. Increasing complexity therefore improved
the simple baseline more consistently than it improved a transparent temporal
model.

The bounded official-architecture run completed through the official Sensorium
loader and model factory for five mice. Correlations ranged from 0.011 to 0.035
under the deliberately restricted training budget. The artifact establishes
software interoperability, not competitive model quality. The audit accordingly
marks it as a bounded integration control and blocks state-of-the-art or
leaderboard-equivalent interpretations.

\subsection{The aggregate MICRONS analysis prevents an initial overclaim}

The first expanded MICRONS analysis compared connected pairs with random,
distance-matched, and degree-matched non-connected pairs at aggregate level.
Connected pairs were more functionally similar than random pairs and
degree-matched controls, but the contrast did not survive distance matching.
Because cortical distance influences both connection probability and functional
organization, the aggregate result was classified as negative or inconclusive.

This failure altered the analysis rather than being discarded. It showed that a
random null was too permissive and that the aggregate association could not be
separated from spatial embedding. The expanded result was retained as a stress
test, and subsequent analyses were stratified by functional descriptor and
anatomical context. Readout-location similarity was fixed as the primary
endpoint for the replication package after this diagnostic stage.

Had only the random or degree-matched contrast been reported, the same data
could have supported an overstated positive conclusion. Requiring the
distance-matched gate changed the outcome and forced a narrower hypothesis. This
is a direct example of the claim-audit logic affecting scientific inference
rather than merely relabelling a result after analysis.

\subsection{MICRONS yields a replicated local structure--function association}

The three MICRONS cohorts jointly contain 2991 units, 6575 loaded synapses, and
5943 unique connected directed pairs. Discovery and the two hold-outs contain
2095, 1926, and 1922 connected pairs among 999000, 983072, and 997002 candidate
directed non-self pairs. The primary endpoint passed both distance- and
degree-matched tests in every cohort (Table~\ref{tab:microns-primary}).

\begin{table}[t]
\centering
\caption{Replicated MICRONS primary endpoint: connected directed pairs show closer readout-location similarity than matched non-connected controls.}
\label{tab:microns-primary}
\small
\begin{tabular}{lrrrrrr}
\toprule
Cohort & Units & Connected & $\Delta_d$ & $q_d$ & $\Delta_k$ & $q_k$ \\
\midrule
Discovery & 1000 & 2095 & 0.0215 & 0.0076 & 0.0403 & 0.0053 \\
Hold-out 1000 & 992 & 1926 & 0.0210 & 0.0070 & 0.0380 & 0.0062 \\
Hold-out 2000 & 999 & 1922 & 0.0179 & 0.0087 & 0.0286 & 0.0083 \\
\bottomrule
\end{tabular}
\end{table}

For distance-matched controls, connected-pair advantages in readout-location
similarity were 0.0215, 0.0210, and 0.0179, with one-sided
false-discovery-rate-adjusted $q$ values of 0.0076, 0.0070, and 0.0087.
Degree-matched effects were 0.0403, 0.0380, and 0.0286, with adjusted $q$
values of 0.0053, 0.0062, and 0.0083. The effect retained its sign and
approximate magnitude outside the discovery window.

The discovery family contained 174 statistical tests across functional
descriptors and strata, of which 28 were confirmed positive after
false-discovery-rate correction. The two hold-outs contained 162 and 174 tests,
with 30 and 25 confirmed positives. These counts describe the multiplicity
family and are not presented as independent discoveries. The manuscript claim
rests on the fixed readout-location endpoint and its replication. Bootstrap
stability is summarized in Table~\ref{tab:microns-bootstrap}.

\begin{table}[t]
\centering
\caption{Unit-cluster bootstrap stability for the MICRONS primary endpoint. All lower bounds are positive.}
\label{tab:microns-bootstrap}
\small
\begin{tabular}{lrrrr}
\toprule
Cohort & Distance median & Distance CI95 & Degree median & Degree CI95 \\
\midrule
Discovery & 0.0218 & [0.0166, 0.0263] & 0.0405 & [0.0349, 0.0454] \\
Hold-out 1000 & 0.0209 & [0.0142, 0.0263] & 0.0377 & [0.0321, 0.0431] \\
Hold-out 2000 & 0.0181 & [0.0115, 0.0238] & 0.0287 & [0.0229, 0.0354] \\
\bottomrule
\end{tabular}
\end{table}

Unit-cluster bootstrap intervals remained strictly positive. Distance-matched
lower bounds were 0.0166, 0.0142, and 0.0115; degree-matched lower bounds were
0.0349, 0.0321, and 0.0229. Resampling units rather than pairs reduces the false
precision that would result from treating pairwise observations sharing neurons
as independent. All 300 requested replicates were usable in each cohort.

The result is narrower than a causal mechanism. Readout location is a functional
descriptor associated with connectivity after two major controls, but unmeasured
cell type, layer, area, morphology, or developmental organization may
contribute. The analysis does not perturb edges or establish necessity.

Nor is the association a new biological wiring rule. Ding et al. reported a
substantially richer MICRONS analysis of like-to-like connectivity using
receptive-field, feature, and axon--dendrite geometry controls
\citep{ding2025wiring}. Agreement with that work provides an external-validity
case for MouseBrainBench. The potential novelty lies in applying one explicit
claim-control protocol across synthetic, Allen, Sensorium, and MICRONS regimes,
not in priority over the underlying biological phenomenon.

\subsection{The publication freeze preserves different evidential outcomes}

The final publication artifact admitted separation of predictive, reliability,
out-of-distribution, and mechanistic evidence; Allen as a reproducible negative
mechanistic case; Sensorium as predictive evidence; partial reliability and
topographic evidence in static Sensorium; bounded execution of the official
Sensorium stack; and a replicated local MICRONS readout-location association.

The same artifact blocked a complete mouse-brain digital twin, a
state-of-the-art Sensorium predictor, causal identifiability in Dynamic
Sensorium, a causal interpretation of the MICRONS association, and any
whole-brain extrapolation from the local volume. Replication strengthens the
local observational result but does not remove these restrictions.

The cross-regime result is therefore not that one model achieved the highest
score. It is that the same audit represented three outcomes without semantic
promotion: reproducible but non-identifiable, predictive but non-mechanistic,
and locally structure--function positive but observational. These distinctions
form the empirical support for the methodological contribution.

\section{Discussion}

MouseBrainBench addresses a narrow but consequential problem: computational
neuroscience increasingly produces models that predict neural measurements,
while the language used to describe those models often extends to biological
mechanism or digital-twin fidelity. The experiments show that these
interpretations can be separated operationally. Known-truth systems expose
which evidential component is absent, Allen preserves a strong reproducibility
result while rejecting topology and direction, Sensorium preserves predictive
utility while rejecting causal interpretation, and MICRONS admits a local
structure--function association after matched controls and replication.

The negative Allen result is scientifically useful. A less restrictive
benchmark might report cross-mouse correlation of 0.891 and split-half
correlation of 0.990 as validation of a model. MouseBrainBench instead asks
whether the Allen-informed topology performs better than alternatives and
whether a directed signature can be recovered. It does not. This outcome does
not contradict the biological organization of the visual system. It identifies
an observational limitation: repeatable population dynamics do not uniquely
determine the graph that generated them.

Sensorium illustrates a related distinction in modern Artificial Intelligence.
Static stimulus and context models improve prediction over mean and scrambled
controls, while temporal models and a compact neural network improve predictions
for Dynamic Sensorium. These are valid functional results. They are not evidence
that the architecture's internal variables correspond to synapses, cell types,
or causal operations in the animal. Treating predictive and mechanistic scores
as separate outputs allows the same model to be accepted for one purpose and
rejected for another without contradiction.

The bounded official Sensorium experiment also clarifies what software
integration demonstrates. Executing an official loader and architecture verifies
interoperability, but a restricted local training budget is not comparable with
a published competition result. The very low bounded correlations are therefore
neither evidence against the official architecture nor a competitive baseline.
They document that the integration path works and that a full-budget comparison
remains absent. This prevents an engineering smoke test from becoming an
inflated empirical claim.

The MICRONS result is the strongest positive outcome, but its novelty must be
stated precisely. Ding et al. already demonstrated like-to-like connectivity
with a more extensive biological analysis, including feature, receptive-field,
and axon--dendrite geometry controls. MouseBrainBench does not improve that
biological finding. Instead, the MICRONS experiment tests whether a compact,
predefined audit can recover a compatible local association and preserve its
proper scope. Replication across two non-overlapping windows and positive
unit-cluster bootstrap intervals provide external validity for the audit
pipeline.

The effect is not causal. Readout-location similarity can covary with cortical
area, layer, cell class, developmental organization, or unmeasured morphology.
Distance and degree matching address two major confounds but do not exhaust the
causal graph. In particular, the analysis does not compare axon--dendrite
opportunity at the resolution used by Ding et al., manipulate individual
connections, or demonstrate that the observed edges are necessary for the
functional descriptor. The permitted interpretation is therefore a replicated
local observational association.

The current MIS should also not be treated as a universal scalar definition of
mechanistic explanation. Its principal contribution is structural: evidence
blocks are non-interchangeable and passage criteria are explicit. The particular
criteria and thresholds remain claim-dependent. A molecular mechanism, a
neural-mass model, and an individual-level behavioural twin would require
different blocks. External preregistration and application by independent groups
are needed before MIS can be regarded as a community standard.

Conjunctive rules have an intentional cost. They reduce false promotion of
partial evidence but can reject scientifically interesting models when one
required measurement is unavailable. MouseBrainBench handles this by retaining
block scores and partial-positive labels. Failure of the global decision means
that the target claim is unsupported under the declared evidence; it does not
mean that the model, data, or partial result is useless. This semantic discipline
is important if the framework is to encourage stronger experiments rather than
merely produce negative labels.

The sensitivity analysis indicates that the main qualitative conclusions are
not artifacts of one exact threshold. Allen remains negative under 25\%
relaxation and tightening. Static Sensorium retains partial reliability and
topographic evidence for six of nine threshold combinations, failing when the
reliability threshold reaches 0.70. Dynamic Sensorium gains over mean response
are more stable than gains over temporal SVD, supporting predictive utility but
not superiority of a specific neural architecture. These patterns make the
limitations visible instead of hiding them in a single aggregate score.

Several aspects constrain generalization. MICRONS data originate from one mouse
and a restricted visual-cortical volume. Its three cohorts are independent
windows of units, not independent animals. Allen analyses use a controlled
subset of available sessions. Dynamic Sensorium contains five current and five
legacy mice in the local package. The study contains no intervention, no
behavioural closed loop, and no longitudinal calibration of an individual
animal. Consequently, it cannot support whole-brain, behavioural, causal, or
consciousness claims.

For an engineering contribution in artificial intelligence, the value of the
work depends on the framework being useful beyond this exact collection of
analyses. The software interfaces are therefore organized around claims,
evidence blocks, graphs, predictions, perturbations, and provenance rather than
around hard-coded dataset names. External BMTK, SONATA, or The Virtual Brain
simulations can be added as candidate models without reimplementing their
numerical engines. New datasets can add evidence blocks while retaining the
same publication-freeze contract.

The next scientific step is not to increase model complexity indiscriminately.
It is to add evidence unavailable here: independent-animal structure--function
replication, perturbations that distinguish competing mechanisms, and
full-budget predictive baselines. Those additions would test whether the audit
can move from observational association towards intervention-supported
mechanism. Until then, MouseBrainBench should be understood as a reproducible
claim-control framework for partial mouse-brain models, not as the mouse-brain
model itself.

\section{Conclusions}

This paper presented MouseBrainBench, a claim-aware verification and validation
framework for artificial intelligence models used as partial digital
representations of the mouse brain. The motivating engineering problem is that
predictive performance, reproducibility, structural fidelity, and mechanistic
identifiability are often evaluated with metrics that are subsequently treated
as interchangeable. MouseBrainBench replaces that implicit promotion with
explicit evidence blocks, matched controls, and non-compensatory claim gates.

The Mechanistic Identifiability Score provides a diagnostic summary but does not
decide the claim by weighted averaging. A claim passes only when all required
blocks pass. Synthetic systems demonstrated the intended semantics: common
drive, topology without direction, and direction without topology specificity
were rejected despite strong partial evidence. This behavior is central to the
AI contribution because it turns validation assumptions into executable and
falsifiable contracts.

The real-data experiments produced three different outcomes. Allen Visual
Behavior Neuropixels was highly reproducible but did not identify the proposed
topology or direction. Static and Dynamic Sensorium supported predictive
improvements, including gains from temporal and neural-network models, without
providing mechanistic evidence. MICRONS supplied the positive external-validity
case: connected directed pairs showed greater readout-location similarity than
distance- and degree-matched non-connected pairs in discovery and two
non-overlapping hold-outs. Unit-cluster bootstrap intervals remained positive
in all cohorts.

The MICRONS result is local and observational. It is compatible with more
comprehensive published analyses and is not claimed as a new biological wiring
rule. Its role is to show that the framework can admit a positive claim when a
specific endpoint survives relevant controls and replication. Conversely, the
publication freeze blocks causal, behavioural, whole-brain, consciousness, and
state-of-the-art predictor claims.

MouseBrainBench is therefore not another attempt to emulate an entire mouse
brain. It is an engineering layer for evaluating models that make progressively
stronger claims about complex scientific systems. The next development stage
should add intervention-based datasets, independent-animal structure--function
replication, full-budget external predictors, and adapters for established
multiscale simulators. Those extensions will test whether the same audit can
move from observational association to intervention-supported mechanism without
relaxing its claim boundaries.

\section{Data and code availability}
Source code, configurations, acquisition scripts, and result artifacts are maintained at \url{https://github.com/Fiutten/Mouse-brain}. A versioned archival DOI will be created for the submission release. Raw public datasets are not redistributed. Data acquisition
commands preserve public source identifiers and application programming
interface queries where licenses permit. MICRONS CAVE access requires a
personal token and acceptance of the service terms.

\section{Declaration of competing interest}
The author declares no known competing financial interests or personal
relationships that could have appeared to influence the work reported in this
paper.

\section{Acknowledgements}
The author acknowledges the Allen Institute, the Sensorium and Dynamic
Sensorium teams, the MICRONS Consortium, and the CAVE infrastructure for making
the resources used in this study accessible.

\printcredits
\bibliographystyle{cas-model2-names}
\bibliography{references}


\end{document}
